---
title: "TreeData"
subtitle: "Forest Inventories Harmonization & Correction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TreeData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", echo = T, message = T, warning = F, cache = T
)
```

# Load TreeData and datasets
**Install TreeData**
```{r, eval=F}
devtools::install_github("VincyaneBadouard/TreeData", build_vignettes = TRUE)
```
**Load the packages**
```{r setup}
library(TreeData)
library(knitr)
library(kableExtra)
library(shiny)
library(shinydashboard)
library(shinyjs)
library(shinyWidgets)

# library(ggplot2)
```

# Open the shiny app
```{r, eval = F}
# TreeData::RunApp() # doest' work, why?
shiny::runGitHub( "VincyaneBadouard/TreeData", subdir = "inst/app") # data.tree, stringdist
```


**Load the exemple dataset stored in the package**
```{r}
# data()
```

# All the functions of the package
# Arguments presentation
# Harmonise and correct your data
## Results summary
# Only harmonization

# Decomposed corrections
## Botanical identification correction
## Life status

## Diameter correction
## Recruitment

# Interesting other variables
## Mortality rate
## Recruitment rate
## Growth rate
## Basal area
## Wood density
## Biomass
## Tree height
## Disturbance intensity
# Illustrations
## Diagnostic graphs
## Results graphs

