---
title: "TreeData"
subtitle: "Forest Inventories Harmonization & Correction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TreeData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", echo = T, message = T, warning = F, cache = T
)
```

# Load TreeData and datasets
**Install TreeData**
```{r, eval=F}
devtools::install_github("VincyaneBadouard/TreeData", build_vignettes = TRUE)
```
**Load the packages**
```{r setup}
library(TreeData)
library(knitr)
library(kableExtra)
library(shiny)
library(shinydashboard)
library(shinyjs)
library(shinyWidgets)

# library(ggplot2)
```

# Open the shiny app
```{r, eval = F}
# TreeData::RunApp() # doest' work, why?
shiny::runGitHub( "VincyaneBadouard/TreeData", subdir = "inst/app") # data.tree, stringdist
```


**Load the example dataset stored in the package**
```{r}
# data()
```

# Metadata of the output columns of the TreeData package

- *Comment* (character): error type informations  

**Bota**  
- *Family_TreeDataCor* (character): corrected *Family* name  
- *FamilyCorSource* (character): source of the *Family* correction  
- *Genus_TreeDataCor* (character): corrected *Genus* name  
- *Species_TreeDataCor* (character): corrected *Species* name  
- *BotanicalCorrectionSource* (character): source of the *Genus* and *Species*
correction  
- *ScientificName_TreeDataCor* (character): corrected *ScientificName*  
- *VernName_TreeDataCor* (character): completed if information available at *IdTree*
level. 

**Life Status**  
- *LifeStatus_TreeDataCor* (logical): corrected stem life status.  

**Taper**  
- *TaperDBH_TreeDataCor* (numeric): corrected trees diameter at default *HOM*  

**Diameter**  
- *Diameter_TreeDataCor* (numeric): corrected trees diameter at default *HOM*  
- *DiameterCorrectionMeth* (character): diameter correction methode = "*taper*"/"*local linear regression*"/"*weighted mean*"
/phylogenetic hierarchical("*species*"/"*genus*"/"*family*"/"*stand*")/"*shift realignment*"/"*Same value*".  
- *POM_TreeDataCor* (factor): *POM* value at which the corrected diameters are proposed. Corresponds to the 1st *POM* value at which the stem was measured  
- *HOM_TreeDataCor* (numeric): *HOM* value at which the corrected diameters are proposed. Corresponds to the *DefaultHOM* if *Taper* correction applied. If not, corresponds to the 1st *HOM* value at which the stem was measured  

**Recruitment**  
- *CorrectedRecruit* (logical): TRUE: the row was added to represent the year when the stem was larger than the minimum diameter, but absent from the dataset. FALSE: original dataset row.  

# All the functions of the package
# Arguments presentation
# Harmonise and correct your data
## Results summary
# Only harmonization

# Decomposed corrections

## General errors detection
Detect errors
 - Remove **duplicated rows**
 - Check **missing value** in
X-YTreeUTM/PlotArea/Plot/Subplot/Year/TreeFieldNum/
IdTree/IdStem/Diameter/POM/HOM/Family/Genus/Species/VernName
 - Check **missing value** (NA/0) in the measurement variables: "Diameter",
"HOM", "TreeHeight", "StemHeight"
 - Check of the **unique association of the IdTree with plot, subplot**
**and TreeFieldNum** (at the site scale)
 - Check **duplicated IdTree/IdStem** in a census (at the site scale)
 - Check for trees **outside the subplot** (not implemented yet)
 - Check **invariant coordinates per IdTree/IdStem**
 - Check **fix Plot and Subplot number** (not implemented yet)

## Botanical identification correction

 - No special characters (typography)
 - No family name in the Genus and Species columns (the suffix "aceae" is
     specific to the family name.
 - Correct spelling of botanical names (*Taxonstand or WorldFlora*)
 - Family & Scientific names match (*BIOMASS::getTaxonomy or WorldFlora*)
 - Update the scientific botanical names with the current phylogenetic
     classification
 - Check **invariant botanical informations per IdTree** (1 IdTree = 1 family,
     1 scientific and 1 vernacular name)

## Life status
 - if UseSize argument chosen : if Diameter != NA -> Alive
 If (the value in bold is modified by the value given after the arrow):
 (the ">" gives the chronological order of the sequence)
 
 - *Dead* > Alive -> NA
 - add rows for the forgotten censuses between 2 'Alive' if chosen
 - Alive > *Dead*/*NA* > Alive -> Alive
 - Alive > *NA* > Dead -> NA
 - Alive > *Dead* > NA -> Dead

 - Alive > *NA* > *NA*:
   if DeathConfirmation > unseens -> NA
   if DeathConfirmation =< unseens -> Dead
   
## Taper correction
Transform the tree diameter measured at a given height into the diameter
corresponding to the default measurement height (HOM), using an allometry.


## Diameter correction


## Recruitment

# Interesting other variables
## Mortality rate
## Recruitment rate
## Growth rate
## Basal area
## Wood density
## Biomass
## Tree height
## Disturbance intensity
# Illustrations
## Diagnostic graphs

 - LifeStatusCorrectionPlot() : plot life status correction result
 - DiameterCorrectionPlot() : plot diameter correction result

## Results graphs

