---
title: "TreeData"
subtitle: "Forest Inventories Harmonization & Correction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TreeData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", echo = T, message = T, warning = F, cache = T
)
```

# Load TreeData and datasets
**Install TreeData**
```{r, eval=F}
install.packages("TreeData")
```
**Load the package**
```{r setup}
library(TreeData)
library(knitr)
library(kableExtra)
# library(ggplot2)
```
**Load the exemple dataset stored in the package**
```{r}
# data()
```

# All the functions of the package
# Arguments presentation
# Harmonise and correct in 1 function
## Results summary
# Only harmonization

# Decomposed corrections
## Botanical identification correction
## Life status
## Diameter correction
## Recruitment

# Interesting other variables
## Mortality rate
## Recruitment rate
## Growth rate
## Basal area
## Wood density
## Biomass
## Tree height
## Disturbance intensity
# Illustrations
