---
title: "BotaCorrection with vernabota"
author: "Vincyane Badouard"
date: "14/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data to fill - Preparation
Colonnes nécessaires : "BotaCertainty" (juste mise en factor dans le code), "BotaSource"

(Créer le cas où BotaSource n'existe pas (on fait avec ce qu'il y a))
+ **vernacular names in lower case** (consistency with the prior))
+ **remove the botanical names when BotaSource = Vern** (on ne veut pas garder des id bota qui ne sont que déduites d'un nom vern. On veut le faire avec VernaBota)
+ *remove subspecies and variety name*
+ generate column GenSp : Genus-species
```{r}
Data <- PrepData(Data) # return a data.table
```

# Create prior (à fournir par l'utilisateur)
(voir si je peux aider l'utilisateur à la créer)
table des associations verna-bota codées "0" ou "1"
- vernacular names in columns (consistency with the data to fill)
- botanical names (Family-Genus-Species) in rows

Family        Genus         Species PresentInGuyaFor acacia franc acajou de guyane adugue aganananga aganiamai
1           Fabaceae      Abarema         jupunba             TRUE            0                0      0          0         0
2   Chrysobalanaceae        Acioa      guianensis            FALSE            0                0      0          0         0
3    Melastomataceae      Aciotis    purpurascens            FALSE            0                0      0          0         0

## Clean the user prior
(Voir ce que je veux garder de cette fct)
+ RemoveIndetSp : *remove the botanical names no fully determined* (FALSE : permettre l'asso Vern-Genus/Family.indet. (AssoByGenus/Fam))
+ *remove botanical names not in Guyafor* (sp presentes ds le prior mais pas dans le data (redemander si c bien ça), sinon donnerait une asso non guyafor sur le data guyafor) (la col est rm à la fin)
+ *remove columns with no association* (only "0") (où?)
Les NA ne sont pas permis dans les colonnes Family-Genus-Species-PresentInGuyaFor
```{r}
Prior <- PrepPrior(Prior, RemoveIndetSp = TRUE, RemoveNotGuyafor = TRUE)
```

# Observation data
*Utiliser les données de l'utilisateur (même site), des autres plots que celui qu'on corrige* (to avoid diversity underestimating)
```{r}
ObsData <- PrepData(ObsData) # return a data.table
```


# Create Alpha
Alpha = posteriors = prior (optional) + obs
```{r}
Alpha <- CreateAlpha()
```


# Bota correction
1 même correction par idtree -> traitement sur une table 1row/ind
```{r}
DataTree <- unique(Data[,list(idTree, VernName, Family, Genus, Species, GenSp)])
```

+ GensSpCor
+ BotaCorCode 
```{r}
DataTree <- cbind(DataTree, GensSpCor = as.factor(NA), BotaCorCode = as.factor(NA))
```

## Already fully determined 
```{r}
DataTree[Species != "Indet.", GensSpCor := GenSp] # no need for replacement, give GenSp name
DataTree[Species != "Indet.", BotaCorCode := "fullyDet"]
```

## No Species name
### No vernacular name (or a vernacurlar name not in Alpha)
#### Genus name ok

GenSp = Genus-species (ScentificName)

Codage de l'absence d'info :
Species/Genus : *Indet.*
VernName : *-*

```{r}
# No vern name, but a genus name
DataTree[Species == "Indet." & VernName == "-" & Genus != "Indet.", GensSpCor := GenSp] # Genus-Indet.

DataTree[Species == "Indet." & VernName == "-" & Genus != "Indet.", BotaCorCode := "Det2Genus"] # determined to the genus

# A vern name but not in Alpha, but a genus name
DataTree[Species == "Indet." & VernName != "-"  & !(VernName %in% colnames(Alpha)) &
           Genus != "Indet.", GensSpCor := GenSp] # Genus-Indet.

DataTree[Species == "Indet." & VernName != "-"  & !(VernName %in% colnames(Alpha)) &
           Genus != "Indet.", BotaCorCode := "Det2Genus"] # determined to the genus

```

#### No genus name
##### Family name ok
```{r}
# No vern name, but a family name
DataTree[Species == "Indet." & VernName == "-" & Genus == "Indet." & Family != "Indet.",
         GensSpCor := paste(Family, "Indet.", sep="-")] # Family-Indet.

DataTree[Species == "Indet." & VernName == "-" & Genus == "Indet." & Family != "Indet.",
         BotaCorCode := "Det2Fam"]  # determined to the family

# A vern name but not in Alpha, but a family name
DataTree[Species == "Indet." & VernName != "-" & !(VernName %in% colnames(Alpha)) &
           Genus == "Indet." & Family != "Indet.",
         GensSpCor := paste(Family, "Indet.", sep="-")] # Family-Indet.

DataTree[Species == "Indet." & VernName != "-" & !(VernName %in% colnames(Alpha)) &
           Genus == "Indet." & Family != "Indet.",
         BotaCorCode := "Det2Fam"] # determined to the family
```

##### No family name
```{r}
# No vern name
DataTree[Species == "Indet." & VernName == "-" & Genus == "Indet." & Family == "Indet.",
         GensSpCor := GenSp] # Indet.-Indet.

DataTree[Species == "Indet." & VernName == "-" & Genus == "Indet." & Family == "Indet.",
         BotaCorCode := "NoCor"] # No possible correction

# A vern name but not in Alpha
DataTree[Species == "Indet." & VernName != "-" & !(VernName %in% colnames(Alpha)) &
           Genus == "Indet." & Family == "Indet.",
         GensSpCor := GenSp] # Indet.-Indet.

DataTree[Species == "Indet." & VernName != "-" & !(VernName %in% colnames(Alpha)) &
           Genus == "Indet." & Family == "Indet.",
         BotaCorCode := "NoCor"] # No possible correction
```

### Vernacular name ok and present in Alpha)
```{r}
DataTree[Species == "Indet." & VernName != "-" & VernName %in% colnames(Alpha),
         
         c("GensSpCor", "BotaCorCode") := DrawBota(Alpha = Alpha, # posteriors
                                                   VernName2Sim = as.character(VernName), # the vern name (1 tree)
                                                   Family2Sim = as.character(Family), # the fam name (1 tree)
                                                   Genus2Sim = as.character(Genus), # the gen name (1 tree)
                                                   eps = eps, # background noise for sp not associated with a given vern name
                                                   Determ = Determ),
         by = idTree]
```

#### Ce qui se passe dans DrawBota()
##### No species, Vern name ok, Genus ok
Lorsqu'il y a une asso genre-vern dans data2fill qui ne matche pas avec Alpha, on garde l'info data2fill
```{r}
# get AlphaV, a data.frame with the bota info (4 1er colums) and one colum for the tree vern name
AlphaV <- data.table(Alpha[, 1:4, with = FALSE], Alpha[,VernName2Sim, with = FALSE])
colnames(AlphaV)[length(colnames(AlphaV))] <- "alphaV"

# create an empty vector to store the result
res <- list(as.character(NA), as.character(NA))

if (Genus2Sim != "Indet.") { # Genus ok
  # Put 0 to alphaV for bota names associated with this vernacular but with different genus
  AlphaV[alphaV != 0 & Genus != Genus2Sim, alphaV := 0] # bad genus
  # case with no compatibility between the Genus name and the vernname in Alpha
  if (sum(AlphaV$alphaV) == 0) { # no good genus
    res <- list(paste(Genus2Sim, "Indet.", sep = "-"), "Det2Genus") # on garde le Genus de Data2fill
  }else{
    # case with compatibility (same genus-vern asso, alphaV = 1)
    # for same vern name but different genus and for different vernacular name
    # gives a freq = eps/nb (of other asso)
    AlphaV[alphaV == 0, alphaV := eps/dim(AlphaV[alphaV == 0])[1]] # pas compris
    
    if(Determ == TRUE){ # the more likely association
      res <- DrawDeterm(AlphaV, levelAsso = "AssoByGenus", VernName2Sim = VernName2Sim)
    }
    if(Determ == FALSE) { # random sampling in the alpha matrix
      # Draw in Dirichlet-Categorical
      res[1] <- DirichCat(alphaV = AlphaV$alphaV, Names = AlphaV$GenSp, Vern = VernName2Sim)
      res[2] <- "AssoByGenus"
    }
  }
}
```

##### No species, no genus, Vern name ok, Family ok
```{r}
if (Family2Sim!="Indet." & Genus2Sim=="Indet.") {
  # Put 0  to alphaV for bota names associated with this vernacular but of different family
  AlphaV[alphaV!=0 & Family!=Family2Sim, alphaV:=0]
  # case with no compatibility between the Family name and the vernname in Alpha
  if (sum(AlphaV$alphaV)==0) {
    res <- list(paste(Family2Sim, "Indet.", sep="-"), "Det2Fam")
  } else {
    # case with compatibility
    # for same vernacular name but different family And for different vernacular name
    # gives a freq =eps/nb
    AlphaV[alphaV==0, alphaV := eps/dim(AlphaV[alphaV==0])[1]]
    
    if(Determ==TRUE){
      res <- DrawDeterm(AlphaV, levelAsso="AssoByFam", VernName2Sim=VernName2Sim)
    }
    if(Determ==FALSE) {
      # Draw in Dirichlet-Categorical
      res[1] <- DirichCat(alphaV = AlphaV$alphaV, Names = AlphaV$GenSp, Vern = VernName2Sim)
      res[2] <- "AssoByFam"
    }
  }
}
```

##### No species, no genus, no family, Vern name ok
```{r}
if (Family2Sim=="Indet." & Genus2Sim=="Indet.") {
  # gives a freq =eps/nb for different vernacular name
  AlphaV[alphaV==0, alphaV := eps/dim(AlphaV[alphaV==0])[1]]
  if(Determ==TRUE){
    res <- DrawDeterm(AlphaV, levelAsso="AssoByVern", VernName2Sim=VernName2Sim)
  }
  if(Determ==FALSE) {
    # Draw in Dirichlet-Categorical
    res[1] <- DirichCat(alphaV = AlphaV$alphaV, Names = AlphaV$GenSp, Vern = VernName2Sim)
    res[2] <- "AssoByVern"
  }
}
```

#### Ce qui se passe dans DrawDeterm() (the more likely association)
```{r}
# create an empty vector to store the result
res <- list(as.character(NA),as.character(NA))

Spmax <- AlphaV[alphaV==max(alphaV),GenSp] # max de la colonne alphaV : the more likely association
# if no tie
if (length(Spmax)==1) { # 1 seul max
  res[1] <- as.character(Spmax)
  res[2] <- paste(levelAsso, "Determ", sep="")
} else { # several more likely associations
  # if tie
  res[1] <- as.character(sample(x=Spmax, size=1))
  res[2] <- paste(levelAsso, "DetermT", sep="")
  warning(paste("There were several species with the same probability of association for ", VernName2Sim,
                ", the species was chosen randomly on a tree by tree basis"))
}
```


## Copy the correction by idTree
```{r}
DataSim <- merge(Data, DataTree[,list(idTree, GensSpCor, BotaCorCode)], by = "idTree")

```














