---
title: "DiameterCorrection - new version"
author: "Vincyane Badouard"
date: "19/04/2022"
output: html_document
---

# Packages libraries
```{r, include = FALSE}
library(TreeData)
library(data.table)
library(ggplot2)
# importfrom utils::txtProgressBar
```

# Import data & other inputs
```{r}
data("TestData") # import data
# INPUTS
## data.frame to data.table
setDT(TestData) # with "set" "<-" is not necessary

# Args
MaxDBH = 500 # (numeric, 1 value) in cm 
PositiveGrowthThreshold = 5 # (numeric, 1 value) in cm/year : a tree widening by more than x cm/year is considered abnormal
NegativeGrowthThreshold = -2 # (numeric, 1 value) in cm/census : The possible positive measurement error (+n) cannot be corrected until the growth appears abnormal, but a negative measurement error can be allowed until -n (a tree does not decrease). Thus the positive measurement error (+n) is "compensated".

TrustMeasSet = "first" # "first"/"last" (charcater, 1 value)
WhatToCorrect = c("punctual", "POM change", "shift")
CorrectionType = c("taper", "linear", "quadratic", "individual", "phylogenetic hierarchical") # ("taper" if HOM)
TaperParameter = function(DAB, HOM) 0.156 - 0.023 * log(DAB) - 0.021 * log(HOM) # Cushman et al.2021 (function)
TaperFormula = function(DAB, HOM, TaperParameter) DAB / (2 * exp(- TaperParameter*(HOM - 1.3))) # Cushman et al.2021 (function)

DBHRange = 10 # (numeric, 1 value) in cm
MinIndividualNbr = 5

DefaultHOM = 1.3 # (numeric, 1 value) in m
Pioneers = c("Cecropia","Pourouma") # verna or scfic names ? (characters vector)
PioneersGrowthThreshold = 7.5  # (numeric, 1 value) in cm
IgnorePOM = FALSE # (logical) est-ce que je garde ? on le prend s'il existe
DetectOnly = FALSE # (logical)
```

# Create data with different type of diameters error
- 1 punctual error : growth > 5 cm DBH/year
- 1 punctual error : growth < -2 cm DBH/census
- 2 shift error with available POM : growth < -2 cm DBH/census with no return to normal
- 1 shift error without available POM : growth < -2 cm DBH/census with no return to normal
```{r}
# Create data
## With regular time between census
# Data <- data.table(IdTree = c("a", "b", "c", "d"), # 4 ind
#                    Year = rep(seq(2000,2016, by = 2), 4), # 9 DBH values each (2 years between 2 census)
#                    DBH  = c(13:21), # 1 cm/census (0.5 cm/year)
#                    POM = 0) # default POM

## With irregular time between census
Data <- data.table(IdTree = c("a", "b", "c", "d"), # 4 ind
                   ScientificName = c("A", "B", "C", "D"),
                   Genus = c("Ananas", "Banane", "Citron", "Dachine"),
                   Family = c("Bromeliaceae", "Musaceae", "Rutaceae", "Araceae"),
                   Year = rep(c(seq(2000,2008, by = 2), 2012, 2014,2016, 2020), 4), # 9 DBH values each (2 years between 2 census)
                   DBH  = c(13:17, 19:21, 23), # 0.5 cm/year
                   POM = 0, # default POM
                   HOM = 1.3) # default HOM

# check their growth
ggplot(Data) +
  aes(x = Year, y = DBH) +
  geom_line(size = 0.5, colour = "#46337E") +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal() +
  facet_wrap(vars(IdTree), scales = "free")

# Create errors
Data <- Data[order(IdTree, Year)]

## 1 punctual error : growth > 5 cm DBH/year
Data[IdTree %in% "a", DBH := ifelse(Year %in% 2008, 16+(5*2)+4, DBH)]

## 1 punctual error : growth < -2 cm DBH/census
Data[IdTree %in% "b", DBH := ifelse(Year %in% 2008, 16-4, DBH)] 

## 2 punctual errors : growth > 5 cm DBH/year & growth < -2 cm DBH/census
Data[IdTree %in% "a", DBH := ifelse(Year %in% 2014, 19-4, DBH)] 

## 2 shift error with available POM : growth < -2 cm DBH/census with no return to normal

Data[IdTree %in% "c" & Year %in% c(2008:2014), c("POM", "HOM") := list(1, 1.5)]  # POM = 1, HOM = 1.5

Data[IdTree %in% "c" & Year %in% c(2016:2020), c("POM", "HOM") := list(2,2)] 

Data[POM %in% 1, ("DBH") := c(16-4, (16-4)+2, (16-4)+3)] # 1st shift, always 0.5cm/yearr
Data[POM %in% 2, ("DBH") := c(15-4, (15-4)+2)] # 2nd shift, always 0.5cm/year


## 2 shift error without available POM : growth < -2 cm DBH/census with no return to normal
Data[IdTree %in% "d", c("POM", "HOM") := NA] # no POM
Data[IdTree %in% "d" & Year %in% c(2008:2020), ("DBH") := c(16-4, (16-4)+2, (16-4)+3, (16-4)+4, (16-4)+6)] # only 1 shift

# check their growth
ggplot(Data) +
  aes(x = Year, y = DBH) +
  geom_line(size = 0.5, colour = "#46337E") +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal() +
  facet_wrap(vars(IdTree), scales = "free")

# 1 ind at a time
id <- "c"
DataTree <- Data[IdTree %in% id]
IdTree <- unique(Data[IdTree %in% id, IdTree])
DBHCor <- DBH <- Data[IdTree %in% id, DBH]
Time <- Data[IdTree %in% id, Year]
POM <- Data[IdTree %in% id, POM]
HOM <- Data[IdTree %in% id, HOM]
```

# Taper correction 
Pour les arbres mesurés au dessus du POM par défaut, et dont on a pris la hauteur de mesure.

If HOM (demander à Valentine des colonnes distinctes pour des POM codés et vrais HOM)
- create an argument for the taper formula: 'TaperFormula' and its taper parameter 'TaperParameter'
- check if the HOM column exists
- add a column 'DiameterCorrectionMeth'

- Cushman taper formula without wood density: 
*TaperParameter* : b = 0.156 - 0.023 * log(DAB) - 0.021 * log(HOM)
*TaperFormula* : DAB = EDBH * (2) * e^(- b(HOM - DefaultHOM))
=>  EDBH =  DAB / (2 * e^(- b(HOM - DefaultHOM)))

b : predict taper (unitless) (per individual)

DAB : Diameter Above Buttress (Size, d) in cm

HOM : Height Of Measurement (in m)

EDBH : Equivalent Diameter at 1.3 m

```{r}
# check if the HOM column exists
if(!"HOM" %in% names(Data)){
  stop("you have chosen to make a 'taper' correction,
       but you do not have the necessary 'HOM' column in your dataset")
  
} 

if(any(DataTree[,HOM] > DefaultHOM)){ # if some measurements of the tree were made above the POM by default
  
  DataTree <- GenerateComment(DataTree,
                              condition = (DataTree[,HOM] > DefaultHOM),
                              comment = paste0("HOM different from the default HOM"))
  
  if(DetectOnly %in% FALSE){
    # Apply taper correction
    DataTree[HOM == DefaultHOM, DBHCor := ifelse(is.na(DBHCor), DBH, DBHCor)] # At default POM, keep the measured value
    DataTree[HOM > DefaultHOM, DBHCor := round(TaperFormula(DAB = DBH,
                                                            HOM = HOM,
                                                            TaperParameter = TaperParameter(DAB = DBH, HOM = HOM)),
                                               digits = 1)
    ]
    
    # Add the column with the correction method
    DataTree[HOM > DefaultHOM & !is.na(DBHCor), DiameterCorrectionMeth := "taper"]
    
    
  }
}

```

# Punctual error correction

## Punctual error detection
```{r}
# 1 ind at a time
id <- "a"
IdTree <- unique(Data[IdTree %in% id, IdTree])
DBHCor <- DBH <- Data[IdTree %in% id, DBH]
Time <- Data[IdTree %in% id, Year]
POM <- Data[IdTree %in% id, POM]

# Compute diameter incrementation -----------------------------------------------------------------------------------------
cresc <- ComputeIncrementation(Var = DBHCor, Type = "annual", Time = Time)
cresc_abs <- ComputeIncrementation(Var = DBHCor, Type = "absolute", Time = Time)

# Detect abnormal growth --------------------------------------------------------------------------------------------------      
Ncresc_abn <- sum(cresc >= PositiveGrowthThreshold | cresc_abs < NegativeGrowthThreshold) # nbr of abnormal values
# le retour à la normale est considéré comme une erreur (perte excessive)

if(Ncresc_abn > 0) { # if there are abnormal values
  for (i in 1:Ncresc_abn) { # for each abnormal value
    
    # Begin with the census with the highest DBH change
    cresc_abn <- which(cresc >= PositiveGrowthThreshold | cresc_abs < NegativeGrowthThreshold) # quels sont les concernés
    ab <- cresc_abn[which.max(abs(cresc[cresc_abn]))] # the highest absolute DBH increment (celui qui a initié) (ab in cresc indices)
    
    # Check if there is a return to normal --------------------------------------------------------------------------------
    if(length(ab) == 1) {
      # With 4 values surrounding ab
      surround <- c(ab - 2, ab - 1, ab + 1, ab + 2) # the 4 values, 2 before & 2 after the error
      # In the DBH seq
      surround <- surround[surround > 0 &
                             surround <= length(cresc)] # de taille maximale [0;longueur de l'incrémentation = length(dbh) -1)]
      
      # Moment of max and min DBH changes around ab (including ab, that should be one of the 2)
      sourround_ab <- sort(c(surround, ab)) # ab and its 4 values around in the increasing order
      up <- sourround_ab[which.max(cresc[sourround_ab])] # Max growth of this seq
      down <- sourround_ab[which.min(cresc[sourround_ab])] # Min growth of this seq
      
      if(length(surround) > 0) {  # if there are values around ab
        # 1st case = Punctual: excessive increase/decrease offset by a similar decrease in dbh, + a normal growth
        # is there a value that could compensate the excessive DBH change?
        # check if removing those values would solve the problem (ie cresc < 5 & cresc_abs > -2 )
        if( # if it's punctual --------------------------------------------------------------------------------------------
            # the max positive growth is before the min negative growth (increase then decrease)
            isTRUE(down > up & cresc[up] * cresc[down] < 0 &
                   # Compute cresc around the error by skipping the error to check if it's normal
                   # (cresc and down are cresc indices, to have the corresponding DBH index add +1)
                   ((DBHCor[down + 1] - DBHCor[up]) / (Time[down + 1] - Time[up])) <= PositiveGrowthThreshold &
                   (DBHCor[down + 1] - DBHCor[up]) >= NegativeGrowthThreshold) | # Nino a mis un else plutot qu'un "ou" (ce qui semble mieux)
            # the max positive growth is after the min negative growth (decrease then increase)
            isTRUE(up > down & cresc[up] * cresc[down] < 0 &
                   # Compute cresc around the error by skipping the error to check if it's normal
                   ((DBHCor[up + 1] - DBHCor[down]) / (Time[up + 1] - Time[down])) <= PositiveGrowthThreshold &
                   (DBHCor[up + 1] - DBHCor[down]) >= -NegativeGrowthThreshold)) { # different de Nino (tag dirtyhack) (runner ac mon ex)
          
          # Abnormal DBH <- NA and will be replaced later on (by RegressionInterpolation()) -------------------------------
          first <- min(up, down) + 1 # The punctual error: the 1st value with the greatest increment (positive or negative) (+1 to switch from cresc to DBH indices)
          last <- max(up, down) # The compensation: the last value with the greatest increment (positive or negative) (in cresc indices, -1 to get the DBH index)
          
          # first and last is the same value if the error is compensated immediately
          DBHCor[first:last] <- NA # put NA from the 1st to the last greatest increment
        } # if it's a punctual error, else it's a shift
        
      } # if there are values around ab
      
      ## If only 2 values, with abnormal difference
      if(length(DBHCor[!is.na(DBHCor)]) == 2 & i==1){ # i =  chaque valeur aberrante
        
        if(TrustMeasSet %in% "first"){
          # trust the 1st one
          DBHCor[!is.na(DBHCor)][2] <- DBHCor[!is.na(DBHCor)][1]
        }else if(TrustMeasSet %in% "last"){
          # trust the 2nd one
          DBHCor[!is.na(DBHCor)][1] <- DBHCor[!is.na(DBHCor)][2]
        }
      }
      
      # Update diameter incrementation (for the i loop)--------------------------------------------------------------------
      cresc <- ComputeIncrementation(Var = DBHCor, Type = "annual", Time = Time)
      cresc_abs <- ComputeIncrementation(Var = DBHCor, Type = "absolute", Time = Time)
      
    } # length(ab) == 1
  } # i loop end
} # if there are abnormal values
```

# Shift error correction

## Individual correction

## Phylogenetic hieracchical correction
+ *init shift*
- Estimer la valeur que devrait avoir le DBH décalé : EstDBH = valeur précédente + dernier cresc et si pas cresc     PositiveGrowthThreshold
- Species, genus ou family level : Colleagues = n ind[EstDBH -DBHRange/2 ; EstDBH +DBHRange/2] > MinIndividualNbr
- DBH[init shift] = previous value +  mean(cresc[Colleagues] 

+ *le shift* : DBH[shift] = valeur prec + leur cresc

```{r}
#### Estimate the shifted DBH ####
if(length(cresc)>0){
  # EstDBH <- previous value + last cresc
}else{
  # EstDBH <- previous value + stand median cresc in 
  #   if TrustMeasSet = "first" :
  # [HealthyValue ; MaxDBH]
  # 
  # or if TrustMeasSet = "last" :
  # [MinDBH ; HealthyValue]
}

#### Find colleagues ####
# Colleagues = n ind[EstDBH -DBHRange/2 ; EstDBH +DBHRange/2] > MinIndividualNbr


#### Correct init shift ####
# DBH[init shift] =  previous value + mean(Colleaguescresc)

#### Correct the shift ####
# DBH[shift] = previous value + their cresc
```



