---
title: "DiameterCorrection"
author: "Vincyane Badouard"
date: "16/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Inspiration: adaptation from Camille Piponiot's and Nino Page (ForestData package) original correction functions.

Return: Create the 'DBHCor' column and fill the 'Comment' column
et col pour le type de cor ?

DBH : diameter in cm

Tests :
- pas de croissance annuelle aberrante > PositiveGrowthThreshold
- pas de décroissance aberrante < NegativeGrowthThreshold

# Packages libraries
```{r, include = FALSE}
library(TreeData)
library(data.table)
library(ggplot2)
# importfrom utils::txtProgressBar
```

# Import data & other inputs
```{r}
data("TestData") # import data
# INPUTS
## data.frame to data.table
setDT(TestData) # with "set" "<-" is not necessary

# Args
PositiveGrowthThreshold = 5 # (numeric, 1 value) in cm/year : a tree widening by more than x cm/year is considered abnormal
NegativeGrowthThreshold = -2 # (numeric, 1 value) in cm/census : The possible positive measurement error (+n) cannot be corrected until the growth appears abnormal, but a negative measurement error can be allowed until -n (a tree does not decrease). Thus the positive measurement error (+n) is "compensated".
DefaultPOM = 1.3 # (numeric, 1 value) in m
Pioneers = c("Cecropia","Pourouma") # verna or scfic names ? (characters vector)
PioneersTreshold = 7.5  # (numeric, 1 value) in cm
IgnorePOM = FALSE # (logical) est-ce que je garde ? on le prend s'il existe
DetectOnly = FALSE # (logical)

```

# Create data with different type of diameters error
- 1 punctual error : growth > 5 cm DBH/year
- 1 punctual error : growth < -2 cm DBH/census
- 2 shift error with available POM : growth < -2 cm DBH/census with no return to normal
- 1 shift error without available POM : growth < -2 cm DBH/census with no return to normal
```{r}
# Create data
Data <- data.table(IdTree = c("a", "b", "c", "d"), # 4 ind
                   CensusYear = rep(seq(2000,2016, by = 2), 4), # 9 DBH values each (2 years between 2 census)
                   DBH  = c(13:21), # 1 cm/census (0.5 cm/year)
                   POM = 0) # default POM

# check their growth
ggplot(Data) +
  aes(x = CensusYear, y = DBH) +
  geom_line(size = 0.5, colour = "#46337E") +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal() +
  facet_wrap(vars(IdTree), scales = "free")

# Create errors
Data <- Data[order(IdTree, CensusYear)]

## 1 punctual error : growth > 5 cm DBH/year
Data[IdTree %in% "a", DBH := ifelse(CensusYear %in% 2008, 16+(5*2)+4, DBH)]

## 1 punctual error : growth < -2 cm DBH/census
Data[IdTree %in% "b", DBH := ifelse(CensusYear %in% 2008, 16-4, DBH)] 

## 2 punctual errors : growth > 5 cm DBH/year & growth < -2 cm DBH/census
Data[IdTree %in% "a", DBH := ifelse(CensusYear %in% 2012, 18-4, DBH)] 

## 2 shift error with available POM : growth < -2 cm DBH/census with no return to normal

Data[IdTree %in% "c" & CensusYear %in% c(2008:2012), ("POM") := 1]  

Data[IdTree %in% "c" & CensusYear %in% c(2014:2016), ("POM") := 2] 

Data[POM %in% 1, ("DBH") := c(16-4, (16-4)+1, (16-4)+2)] # always 1cm/census (0.5cm/year)
Data[POM %in% 2, ("DBH") := c(14-4, (14-4)+1)] # always 1cm/census (0.5cm/year)


## 1 shift error without available POM : growth < -2 cm DBH/census with no return to normal
Data[IdTree %in% "d", ("POM") := NA] # no POM
Data[IdTree %in% "d" & CensusYear %in% c(2008:2016), ("DBH") := c(16-4, (16-4)+1, (16-4)+2, (16-4)+3, (16-4)+4)] # same shift


# check their growth
ggplot(Data) +
  aes(x = CensusYear, y = DBH) +
  geom_line(size = 0.5, colour = "#46337E") +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal() +
  facet_wrap(vars(IdTree), scales = "free")

# 1 ind at a time
id <- "d"
IdTree <- unique(Data[IdTree %in% id, IdTree])
DBHCor <- Data[IdTree %in% id, DBH]
Time <- Data[IdTree %in% id, CensusYear]
POM <- Data[IdTree %in% id, POM]
```

# Ask to the user

# Check args
```{r}
# Data
if (!inherits(Data, c("data.table", "data.frame")))
  stop("Data must be a data.frame or data.table")

# PositiveGrowthThreshold/NegativeGrowthThreshold/PioneersTreshold/DefaultPOM (numeric, 1 value)
if(!all(unlist(lapply(list(PositiveGrowthThreshold, NegativeGrowthThreshold, DefaultPOM, PioneersTreshold),
                      length)) %in% 1) |
   !all(unlist(lapply(list(PositiveGrowthThreshold, NegativeGrowthThreshold, DefaultPOM, PioneersTreshold),
                      inherits, c("numeric", "integer")))))
  stop("The 'PositiveGrowthThreshold', 'NegativeGrowthThreshold', 'PioneersTreshold' and 'DefaultPOM' arguments
         of the 'DiameterCorrection' function must be 1 numeric value each")

# Pioneers (characters vector)
if (!inherits(Pioneers, "character"))
  stop("'Pioneers' argument of the 'DiameterCorrection' function must be a characters vector")

# IgnorePOM/DetectOnly (logical)
if(!all(unlist(lapply(list(IgnorePOM, DetectOnly),
                      inherits, "logical"))))
  stop("The 'IgnorePOM' and 'DetectOnly' arguments
         of the 'DiameterCorrection' function must be logicals")

# DBH column exists
if(!"DBH" %in% names(Data))
  stop("The 'DBH' (Diameter at Breast Height) column does't exist in the dataset")

```

# Errors detection
```{r}
# Missing measure
Vars <- c("DBH", "POM")

for (v in 1:length(Vars)) {
  
  if(Vars[v] %in% names(Data)){ # If the column exists
    
    Data <- GenerateComment(Data,
                            condition = is.na(Data[,get(Vars[v])]),
                            comment = paste0("Missing value in ", Vars[v]))
  }
}

```

# Internal function: Replace incorrect punctual abnormal DBH (missing values in corrected col)
```{r}
# Description : Correction of punctual errors by linear interpolation
# Replace incorrect punctual abnormal DBH (missing values in corrected col)
# Source : Camille Piponiot
# Return : vector of corrected DBHs

# Y = DBHCor
# Time : year
# Status = NULL

LinearInterpolationCorrection <- function(Y, Time) {
  
  miss <- which(is.na(Y)) # DBH = NA -> value to replace
  
  # miss: nb of the missing value(s) (which values) (vector)
  
  Y[miss] <- sapply(miss, function(i) { # i = each value de miss
    
    # case 1: i is the first value of the series
    if (i < min(which(!is.na(Y)))) {
      ## choose 2 next values
      yval <- Y[which(!is.na(Y))[1:min(2, sum(!is.na(Y)))]] # DBH de la 1ere val non-NA à la 2nde si elle existe
      tval <- Time[which(!is.na(Y))[1:min(2, sum(!is.na(Y)))]] # les années concernées de ces 2 val
    }
    
    #case 2: i is the last value of the series
    else if (i > max(which(!is.na(Y)))) {
      yval <- Y[which(!is.na(Y))[(sum(!is.na(Y)) - 1):sum(!is.na(Y))]] # DBH de l'avant-derniere val non-NA à la dernière
      tval <-
        Time[which(!is.na(Y))[(sum(!is.na(Y)) - 1):sum(!is.na(Y))]] # années concernées
      yval <- yval[!is.na(yval)] # s'assurer que ce sont des non-NA
      tval <- tval[!is.na(yval)]
    }
    
    # case 3: i is in the middle of the series
    else {
      yval <-
        Y[c(max(which(!is.na(Y[1:(i - 1)]))), i + min(which(!is.na(Y[(i + 1):length(Y)]))))] # DBH
      # prendre la valeur la plus récente non-NA avant le NA, et la prochaine valeur non-NA après le NA = prendre les valeurs encadrantes non-NA les plus proches du NA
      tval <-
        Time[c(max(which(!is.na(Y[1:(i - 1)]))), i + min(which(!is.na(Y[(i + 1):length(Y)]))))] # années concernées
    }
    
    # Linear regression: DBH ~ years
    reg <- lm(yval ~ tval)$coef # extract the coefs
    yi <- reg[1] + reg[2] * Time[i] # (y = b + ax),  DBHi = b + a YEARi
    
    if (sum(!is.na(tval)) == 1) { # s'il n'y a qu'1 valeur non-anormale
      yi <- yval # le DBH à corriger prends la valeur de l'unique DBH non-anormal
    }
    
    return(yi) # DBH of i, yi -> Y[miss]
    
  }) # sapply end (for each i)
  
  return(unlist(Y)) # corrected DBHs
}

```

# Compute absolute and annual diameter incrementation
```{r}
# Initialisation
cresc <- cresc_abs <- rep(0, length(DBHCor) - 1) # (cresc[1] corresponds to the 2nd DBH)
# cresc_abs : absolute diameter increment (not divided by time between census)
# cresc : annual diameter increment

if (sum(!is.na(DBHCor)) > 1) { # if there is at least 1 measurement
  
  # Annual diameter increment (cresc)
  cresc[which(!is.na(DBHCor))[-1] - 1] <- # 8 cresc for 9 dbh values ([-1]), shift all indices by 1 to the left (-1)
    diff(DBHCor[!is.na(DBHCor)]) / diff(Time[!is.na(DBHCor)]) # DBH difference between pairwise censuses / time difference between pairwise censuses
  
  # Absolute diameter increment (not divided by time between census) (cresc_abs)
  cresc_abs[which(!is.na(DBHCor))[-1] - 1] <- diff(DBHCor[!is.na(DBHCor)]) 
  
}

```

# If the POM exists and varies
0 : default POM
1 : another POM
2 : another POM
```{r}
# je pense qu'il faudrait éviter que qu'il corrige quand le POM est cst (pour les erreurs ponctuelles) (a faire)
if (any(!is.na(POM))) { # POM exists?
  
  raised = which(diff(c(NA, POM)) == 1) # Detection des changements de POM (1ere val = NA car ) (1 = changement de POM)
  
  if (length(raised) == 0 | # if POM is constant
      isTRUE(DBHCor[raised] > DBHCor[raised - 1]) # the 1st DBH at this new POM is higher than the previous DBH (= augmentation du DBH, alors qu'on attend une diminution pour élévation de POM)
  ){
    # là j'ai l'impression qu'elle fait le cas où ya des POM mais n'a pas détécté de changements, donc elle suggère qu'il y eu changement de POM lors de la plus faible incrémentation annuelle (ok), ou le cas où il ya eu augmentation du DBH, alors qu'on attend une diminution (pour élévation de POM), et elle considère tjrs que c'est l'incré min qui doit être corrigée (comprends pas)
    
    raised = which.min(cresc) + 1 # position de la valeur ayant la plus faible de l'incrémentation annuelle (+1 to switch from cresc to DBH indices)
  }
  DBHCor_Vin <- DBHCor
  # rs = 1
  for (rs in 1:length(raised)){ # autant de rs que de changements de POM
    # J'ai fait différent de Camille :
    # j'ai corrigé 1 shift à la fois sans toucher au prochain
    DBHCor_Vin[raised[rs]:min(raised[rs+1]-1, length(DBHCor_Vin), na.rm = TRUE)] <-
      (DBHCor_Vin[raised[rs]:min(raised[rs+1]-1, length(DBHCor_Vin), na.rm = TRUE)]) + # initial DBH values
      (DBHCor_Vin[raised[rs] - 1] - DBHCor_Vin[raised[rs]]) + # lost size
      mean(cresc[((raised[rs]+1):min(raised[rs+1]-1, length(DBHCor_Vin), na.rm = TRUE))-1]) *(Time[raised[rs]+1]-Time[raised[rs]]) # mean of the annual diameter increments in the shift, multiplied by the number of years between the census of the shift and the following one.
    
    # previous version
    # mean(cresc[-((raised[1]:length(DBHCor_Vin))-1)]) *(Time[raised[rs]+1]-Time[raised[rs]]) # mean of the annual diameter increments before the shifts, multiplied by the number of years between the census of the shift and the previous one.
    # (-1 in cresc() to switch from DBH to cresc indices)
    
  }
  
  # Camille version for comparaison
  DBHCor_Cam <- DBHCor
  for (rs in raised) # autant de rs que de changments de POM
    DBHCor_Cam[rs:length(DBHCor_Cam)] = DBHCor_Cam[rs:length(DBHCor_Cam)] + (DBHCor_Cam[rs - 1] - DBHCor_Cam[rs]) + mean(cresc[-(rs -1)])
  
  
  ## Update cresc (for the next corrections)
  cresc <- cresc_abs <- rep(0, length(DBHCor) - 1) # (cresc[1] corresponds to the 2nd DBH)
  
  if (sum(!is.na(DBHCor)) > 1) { # if there is at least 1 measurement
    
    cresc[which(!is.na(DBHCor))[-1] - 1] <- 
      diff(DBHCor[!is.na(DBHCor)]) / diff(Time[!is.na(DBHCor)]) 
    cresc_abs[which(!is.na(DBHCor))[-1] - 1] <- diff(DBHCor[!is.na(DBHCor)]) 
  }
  
}

# Plottons ça 
ggplot() +
  aes(x = Time, y = DBHCor_Cam) +
  geom_line(size = 0.5, colour = "#46337E") +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal()
ggplot() +
  aes(x = Time, y = DBHCor_Vin) +
  geom_line(size = 0.5, colour = "#46337E") +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal()
```

# Punctual error detection and correction
```{r}
####    if there is a DBH change > 5cm/year or < -2 cm   ####
### do as many corrections as there are abnormal DBH change values ###
Ncresc_abn = sum(cresc >= PositiveGrowthThreshold | cresc_abs < NegativeGrowthThreshold) # nbr of abnormal values
# le retour à la normale est considéré comme une erreur (perte excessive)

if (Ncresc_abn > 0) { # if there are abnormal values
  for (i in 1:Ncresc_abn) { # for each abnormal value
    
    # begin with the census with the highest DBH change
    cresc_abn = which(cresc >= PositiveGrowthThreshold | cresc_abs < NegativeGrowthThreshold) # quels sont les concernés
    ab <- cresc_abn[which.max(abs(cresc[cresc_abn]))] # the highest absolute DBH increment (celui qui a initié) (ab in cresc indices)
    
    if (length(ab) == 1) {
      # values surrounding ab
      surround = c(ab - 2, ab - 1, ab + 1, ab + 2) # the 4 values, 2 before & 2 after the error
      # that have a meaning (no NAs or 0 values)
      surround = surround[surround > 0 &
                            surround <= length(cresc)] # de taille maximale [0;longueur de l'incrémentation = length(dbh) -1)]
      
      # moment of max and min DBH changes around ab (including ab, that should be one of the 2)
      sourround_ab = sort(c(surround, ab)) # ab and its 4 values around in the increasing order
      up = sourround_ab[which.max(cresc[sourround_ab])] # l'incrémentation max de cette série
      down = sourround_ab[which.min(cresc[sourround_ab])] # l'incrémentation min de cette série
      
      if (length(surround) > 0) {  # if there are values around ab
        # 1st case : excessive increase/decrease offset by a similar decrease in dbh, plus 5cm/yr
        # is there a value that could compensate the excessive DBH change?
        # check if removing those values would solve the problem (ie cresc < 5 & cresc_abs > -2 )
        if (isTRUE(down > up & cresc[up] * cresc[down] < 0 &
                   # first the most important increase and then the most important decrease of the DBH
                   # cresc and down are cresc indices, to have the corresponding DBH index add +1
                   (DBHCor[down + 1] - DBHCor[up]) / (Time[down + 1] - Time[up]) <= PositiveGrowthThreshold &
                   DBHCor[down + 1] - DBHCor[up] >= NegativeGrowthThreshold) | # Nino a mis un else plutot qu'un "ou" (ce qui semble mieux)
            
            isTRUE(up > down & cresc[up] * cresc[down] < 0 &
                   # first the most important decrease and then the most important increase of the DBH
                   (DBHCor[up + 1] - DBHCor[down]) / (Time[up + 1] - Time[down]) <= PositiveGrowthThreshold &
                   DBHCor[up + 1] - DBHCor[down] >= -NegativeGrowthThreshold)) { # different de Nino (tag dirtyhack) (runner ac mon ex)
          
          # correction: abnormal values are deleted and will be replaced later on (see missing)
          first <- min(up, down) + 1 # the first value with the highest increment (positive or negative) (+1 to switch from cresc to DBH indices)
          last <- max(up, down) # the last value with the highest increment (positive or negative) (in cresc indices, -1 to get the DBH index)
          # first and last is the same value if the error is compensated immediately
          DBHCor[first:last] <- NA # put NA de la 1ere valeur à haut changement, à la dernière valeur à haut changement
        } # if it's a ponctual error
        
      } # if there are values around ab
      
      ## If only 2 values, with abnormal difference
      if (length(DBHCor[!is.na(DBHCor)]) == 2 & i==1){ # i =  chaque valeur aberrante
        # trust the 2nd one
        DBHCor[!is.na(DBHCor)][1] <- DBHCor[!is.na(DBHCor)][2]
      }
      
      # Update cresc (for the i loop)
      cresc <- cresc_abs <- rep(0, length(DBHCor) - 1) # (cresc[1] corresponds to the 2nd DBH)
      
      if (sum(!is.na(DBHCor)) > 1) { # if there is at least 1 measurement
        
        cresc[which(!is.na(DBHCor))[-1] - 1] <- 
          diff(DBHCor[!is.na(DBHCor)]) / diff(Time[!is.na(DBHCor)]) 
        cresc_abs[which(!is.na(DBHCor))[-1] - 1] <- diff(DBHCor[!is.na(DBHCor)]) 
      }
      
    } # length(ab) == 1
  } # i loop end
} # if there are abnormal values

## Replace incorrect punctual abnormal DBH (missing values in corrected col (DBHCor)) by linear interpolation
if (any(!is.na(DBHCor))) {
  
  DBHCor <- LinearInterpolationCorrection(DBHCor, Time) # correction of punctual errors by linear interpolation
  
}else{ # if no measure
  DBHCor = rep(0, length(DBHCor)) # DBH_cor = 0
}

```


# Shift correction without available POM (arrété là)
User choice (TrustMeasSet):
A) Always trust the **1st set** of measurements (ex: because it's at the default POM)
B) Always trust the **last set** of measurements (ex: because new method, safer in your opinion)

C) Trust the set of measurements with **more values** (ex: because safer in your opinion) and:
- (C1) if they are the same size, then we trust the **1st one** (ex: because it's at the default POM)
or
- (C2) if they are the same size, then we trust the **last one** (ex: because new method, safer in your opinion)

Penser qu'il peut y avoir plsrs shifts !
```{r}
TrustMeasSet = "C1"
# Mean DBH change around ab (surround = 4 values max)
MeanLocalcresc <- max(mean(cresc[surround], na.rm = TRUE), 0) # mean(cresc) of these 4 values (without ab). 0 is better than a negative cresc

# else {

# 2 types of correction? (considèrent qu'1 shift possible:
Correc1st <- DBHCor[1:ab] <-
  DBHCor[1:ab] + # inital values
  (DBHCor[ab +1] - DBHCor[ab]) - # on rajoute la valeur gagnée entre l'erreur et sa valeur précédente
  MeanLocalcresc * diff(Time)[ab] # on enlève l'incrémentation moyenne initialement présente * nbr d'années entre l'erreur et la valeur précédente

Correclast <- DBHCor[(ab +1):length(DBHCor)] <-
  DBHCor[(ab +1):length(DBHCor)] - # inital values
  (DBHCor[ab +1] - DBHCor[ab]) + # on enlève la valeur gagnée entre l'erreur et sa valeur précédente
  mean(cresc[-ab]) * diff(Time)[ab] # on ajoute l'incrémentation moyenne sans l'erreur * nbr d'années entre l'erreur et la valeur précédente (Nino utilise MeanLocalcresc)

# Always trust the **1st set** of measurements (ex: because it's at the default POM)
if(TrustMeasSet %in% "A"){
  
  # So, we correct from error to last measurement (ptetre mieux encore une fois de faire shift après shift)
  Correclast
}
# Always trust the **last set** of measurements (ex: because new method, safer in your opinion)
if(TrustMeasSet %in% "B"){
  
  # So, we correct from the 1st measurement to the value before the error
  Correc1st
}
# Trust the set of measurements with **more values** (ex: because safer in your opinion)
if(TrustMeasSet %in% c("C1", "C2")){
  
  # Trust the set of measurements with **more values**
  if ((sum(!is.na(DBHCor[1:ab])) > sum(!is.na(DBHCor))/2)) { # si la serie avant l'erreur est plus longue que le reste des mesures
    
    
  }
  if(TrustMeasSet %in% "C1"){ #  if they are the same size, then we trust the **1st one** (ex: because it's at the default POM)
    Correclast
  }
  
  if(TrustMeasSet %in% "C2"){ # if they are the same size, then we trust the **last one** (ex: because new method, safer in your opinion)
    # So, we correct from the 1st measurement to the value before the error
    Correc1st
  }
  
  
  # Nino fait ça si l'erreur est positive (aug DBH) (et trust the 1st si dim DBH) :
  if ((sum(!is.na(DBHCor[1:ab])) > sum(!is.na(DBHCor))/2)) { # si la serie avant l'erreur est plus longue que le reste des mesures
    
    # On corrige la plus courte
    Correclast
    
  }else{ # si la série après l'erreur est de même longueur ou plus grande que la série avant l'erreur
    # on corrige la 1ere serie (on croit la dernière)
    Correc1st
  }
  
  
}




#   }
# } # if there are values around ab

```







# POM translation (pas encore adapté)
Default POM : 1.3m
Si le POM est diférent en général il est plus élevé (au dessus des contreforts),
mais pourrait être inférieur si on a du éviter un obstacle par ex

```{r}
# cresc = annual size difference

if(!all(POM %in% DefaultPOM)){ # if any POMs are different to the default POM
  
  POM_diff <- diff(POM) # differences between POMs
  
  # Rows with a different POM
  Shifts <- which(!is.na(POM_diff) & POM_diff != 0) 
  
  for(s in Shifts){ # s = rows id with a POM different to the default
    
    existing <- c(s-2, s-1, s+1, s+2) # 2 rows before and 2 after
    existing <- existing[existing > 0 & existing <= length(cresc)] # "0" rows doesn't exist, we can't take it. length(POM) = length(DBH difference)
    existing <- existing[cresc[existing] > negative_growth_threshold &
                           cresc[existing] < positive_growth_threshold] #Because we don't want to use outliers to compute expected growth...
    
    MeanLocalcresc <- max(mean(cresc[existing], na.rm = TRUE), 0) # annual size diff (positive) mean
    
    # Correction from POM
    correction <- max(size_corr[s] - size_corr[s+1], 0) + (MeanLocalcresc*(Time[s+1]-Time[s]))
    # if size increases (max = 0), correction =  (MeanLocalcresc*(Time[s+1]-Time[s]))
    # if size decreases, correction = lost size + (MeanLocalcresc*(Time[s+1]-Time[s]))
    # size evolution (= size at t0 - size at t1) or 0 + annual size diff mean * years nbr between 2 censuses
    
    size_corr[(s+1):length(size_corr)] <- (size_corr[(s+1):length(size_corr)]) + correction
    code_corr[(s+1):length(code_corr)] <- ifelse(code_corr[(s+1):length(code_corr)] == 0, # if no code
                                                 "POM",
                                                 paste(code_corr[(s+1):length(code_corr)], "POM", sep = "+"))
  }
  
}

```


# Errors correction per tree (lire .correct_size_tree)
Si l'arbre est une sp pionnière prendre le threshold approprié
```{r}
if(DetectOnly %in% FALSE){
  
  
}

```

# Errors correction for all trees
```{r}
# Create the DBHCor column (check that it does not already exist) et col pour le type de cor ?

# data.frame to data.table
setDT(Data)

# Tag pioneer sp (T/F) (A FAIRE)
# if VernName %in% Pioneers -> TRUE
if(!PioneerSp %in% names(Data))
  Data[, PioneerSp := ifelse(VernName %in% Pioneers, TRUE, FALSE)]

# Order IdTrees and times in ascending order
Data <- Data[order(IdTree, CensusYear)]

# IdTrees vector
Ids <- as.vector(na.omit(unique(Data$IdTree))) # Tree Ids

# Dataset with the rows without IdTree
DataIDNa <-  Data[is.na(IdTree)]

# Apply for all the trees
pb <- utils::txtProgressBar(min = 0, max = length(Ids), style = 3) # a progress bar per id

# i = "100635"
Data <- do.call(rbind, lapply(Ids, function(i) DiameterCorrectionByTree(
  Data[IdTree %in% i], # per IdTree, all censuses
  Censuses = as.vector(na.omit( # rm NA
    unique(Data[Plot %in% unique(Data[IdTree %in% i, Plot]),  CensusYear]) # the censuses for the plot in which the tree is
  )),
  DetectOnly = DetectOnly
)
)) # do.call apply the 'rbind' to the lapply result

close(pb) # close the progress bar


# Re-put the the rows without IdTree
Data <- rbindlist(list(Data, DataIDNa), use.names=TRUE, fill=TRUE)

```



