---
title: "FieldWay (ErrorsDetection)"
author: "Vincyane Badouard"
date: "29/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objective: Detect errors
- Check of the *unique association of the idTree with plot, subplot*
- Check *duplicated idTree* in a census
- Check *missing value* (NA/NULL/0) of tree size
- Check for *missing coordinates*
- Check for trees *outside the subplot*
- Missing values in other columns (LifeStatus)
- Check *missing trees* compared to the previous census (add rows) (idTree absent of the last census, but alive at the previous census)
- Check *invariant columns*: "Plot","SubPlot","TreeFieldNum","X","Y" for a unique "idTree"

- Bota
- Life status
- Diameter
- Recruitment

-> Create a comment ("Comment" column) to inform about the error type.

# Packages libraries
```{r}
library(TreeData)
library(data.table)
```


# Import data & other inputs
```{r}
data("ParacouSubset") # import data
# INPUTS
Data = ParacouSubset
Plot = "SubPlot"
SubPlot = "SubSubPlot"
Time = c("2016","2017","2018","2019","2020")
TreeFieldNum = "TreeFieldNum"
IdTree = "idTree"
LifeStatus = "CodeAlive"
Size = "Circ"
SizeUnit = "m"
POM = "MeasCode"
POMUnit = "code" # NULL
PlotArea = "PlotArea"
X = "Xutm"
Y = "Yutm"
ScientificName = "none" # NULL
VernName = "none" # NULL
Family = "Family"
Genus = "Genus"
Species = "Species"
CommercialSp = "CommercialSp"
TreeHeight = "none" # NULL
TreeHeightUnit = "none" # NULL
```


# Multiple checks

## Missing values
Plot, SubPlot, Time,TreeFieldNum, IdTree, LifeStatus, Size, POM, X, Y
If the column exists, but have NA values
```{r}
Vars <- c(Plot, SubPlot, Time, TreeFieldNum, IdTree, LifeStatus, Size, POM, X, Y)

for (v in Vars) {
  
  if(Vars[v] %in% names(Data)){ # If the column exists
    
    MissingVal <- Data[is.na(get(Vars[v])) | get(Vars[v]) == 0, get(Vars[v])]
    
    if(any(MissingVal)){ # but have NA values
      Data[MissingVal, Comment := paste0("Missing value in ", Vars[v])] # comments
    }
  }
}

# Simple version
# Check for missing coordinates
# missing_coor <- Data[is.na(get(X)) | is.na(get(Y))] # transformer en 1 vect
# if(any(missing_coor)){
#   
#   Data[missing_coor, Comment := paste0("Missing coordinates: ",X,", ",Y)]
# }
```
## Check of the unique association of the idTree with plot, subplot and TreeFieldNum
```{r}
# Check of the unique association of the idTree with plot, subplot

correspondances <- na.omit(unique(Data[, c(IdTree, Plot, SubPlot, TreeFieldNum), with = FALSE])) # with = FALSE: the column names can be used as variables

CorresIDs <- correspondances[, IdTree, with = FALSE]


if(!identical(CorresIDs, unique(CorresIDs))){ # same lenght, same ids -> 1 asso/ID
  
  Data[duplicated(CorresIDs),
       Comment := paste0("Non-unique association of the ",IdTree," with " ,Plot,", ",SubPlot," and " ,TreeFieldNum)]
  
}


```
## Check duplicated idTree in a census
```{r}
# Check duplicated idTree in a census
ids <- apured_trees$idTree[!is.na(apured_trees$idTree)] # Take care of excluding NAs

if(anyDuplicated(ids) != 0){
  # Extract the duplicated ids - Unique, just in case there is more than two measured for these ids.
  duplicated_ids <- unique(ids[duplicated(ids)])
  
  Data[apured_trees$idTree %in% duplicated_ids, Comment := paste0("Duplicated ",IdTree," in the census")]
  
}
```
## Check for trees outside the subplot
```{r}
# Check for trees outside the subplot

```
## Check missing trees compared to the previous census (add rows)
*Missing tree* = idTree absent of the last census, but alive at the previous census.
```{r}
# Check missing trees compared to the previous census (add rows)
LastTime <- max(Data[,get(Time)]) # Last census year
LastCens <- Data[get(Time) == LastTime]
PreviousTime <- max(unique(Data[get(Time) < LastTime, get(Time)])) # Previous census year
PreviousCens <- Data[get(Time) == PreviousTime]
  
# IdTree's alive at the previous census, in the plot and subplot in common between last and previous census
PreviousCensPlot <- unique(Data[,get(Plot)]) # unique(Data[,c(Plot), with = FALSE])
PreviousCensSubPlot <- 

PrevIDs <- unique(PreviousCens[get(LifeStatus)== TRUE, get(IdTree)])

PrevIDs <- unique(PreviousCens$idTree[PreviousCens$CodeAlive == 1 & PreviousCens$Plot %in% plot & 
                                          PreviousCens$SubPlot %in% subplot])

  unsighted <- PrevIDs[!PrevIDs %in% LastCens$idTree[!is.na(LastCens$idTree)]]
  
  info <- c("Plot","SubPlot","TreeFieldNum","Xfield","Yfield", "idTree")

  if(length(unsighted) > 0){
    unsighted_trees <- PreviousCens[PreviousCens$idTree %in% unsighted & PreviousCens$CodeAlive == TRUE, info]
    unsighted_trees[,names(LastCens)[!names(LastCens) %in% names(unsighted_trees)]] <- NA
    unsighted_trees$to_check <- TRUE
    unsighted_trees$apurement_codes <- "unseen"
    unsighted_trees$comments <- "Arbre oublie/non vu"
    LastCens <- rbind(LastCens,unsighted_trees[,names(LastCens)]) 
  }
  row.names(LastCens) <- NULL
```


## Check invariant columns
```{r}
# "Plot","SubPlot","TreeFieldNum","X","Y"

```

