---
title: "Generate test data"
author: "Vincyane Badouard"
source: "Code adapted from the prep_data_test.R (script_apurement_2020) of Geraldine Derroire"
date: "25/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create errors:
- Non *unique association of the idTree with plot, subplot and TreeFieldNum*
- *duplicated TreeFieldNum* in plot-subplot association
- *duplicated idTree* in a census
- *missing value* (NA/NULL/0) of tree size
- *missing coordinates*
- Trees *outside the subplot*
- Missing values in other columns (Plot, SubPlot, Time, TreeFieldNum, IdTree, LifeStatus, POM)
- *missing trees* compared to the previous census (add rows) (idTree absent of the last census, but alive at the previous census = resurrected tree)
- non-invariant *invariant columns*: "Plot","SubPlot","TreeFieldNum","X","Y" for a unique "idTree"

Internals:
- Bota
- Life status
- Diameter
- Recruitment

# Packages libraries
```{r, include = FALSE}
library(TreeData)
library(data.table)
```

# Import data
```{r}
data("ParacouSubset") # import data
# INPUTS
Data = ParacouSubset
## data.frame to data.table
setDT(Data) # with "set" "<-" is not necessary

nError <- 2 # number of error to add for each case
```

# Missing tree size
```{r}
# Size = NA
modif <- Data[, .I[sample(.N, nError)]] # .I = seq_len(nrow(Data)), .N = nrows in the group -> sample 2 rows number in Data
Data[modif, Circ := NA]
# Data[modif] # to check

# Size = 0
modif <- Data[, .I[sample(.N, nError)]] # .I = seq_len(nrow(Data)), .N = nrows in the group -> sample 2 rows number in Data
Data[modif, Circ := 0]
# Data[modif] # to check
```

# Size with bad precision (pas .0 ou .5)
```{r}
wrong <- c(0.2, 0.3, 0.569, 0.8)
modif <- Data[, .I[sample(.N, nError)]]
Data[modif, Circ := Circ + sample(wrong,1)]
# Data[modif] # to check
```

# Resurrected tree
```{r}
Last_census <- Data[CensusYear == 2020]
Previous_census <- Data[CensusYear == 2019]

# See if a resurrected tree already exists
MortPrev <- Previous_census[CodeAlive == 0 & idTree %in% Last_census[CodeAlive == 1, idTree], idTree]
# dead in 2019, alive in 2020 

Previous_census[idTree == 101410] # dead
Last_census[idTree == 101410] # alive

```

# Missing life status
```{r}
modif <- Data[, .I[sample(.N, nError)]] # .I = seq_len(nrow(Data)), .N = nrows in the group -> sample 2 rows number in Data
Data[modif, CodeAlive := NA]
Data[modif] # to check
```

# Duplicated TreeFieldNum in plot-subplot association
```{r}
modif <- Data[, .I[sample(.N, 1)]] # 1 row to change
duplicatedFieldNum <- Data[!(row.names(Data)) %in% modif & # rows != modif
                             Plot == Data[modif, Plot] & # same plot as modif
                             SubPlot == Data[modif, SubPlot], # same subplot as modif
                           sample(TreeFieldNum,1)] # 1 TreeFieldNum to duplicate

Data[modif, TreeFieldNum := duplicatedFieldNum] # on the row to modif, we duplicate the TreeFieldNum 
# Data[TreeFieldNum == duplicatedFieldNum] # to check
```

# Unseen tree but alive tree after
```{r}

```

# Duplicated idTree in a census
```{r}
idModif <- Last_census[, sample(idTree, nError)] # ne chercher que dans ceux qui ne sont pas à verifier pour eviter de prendre certains avec un duplicatedID # selectionner 2 idTree à modifier

duplicatedID <- Last_census[!(idTree %in% idModif), sample(idTree, 1)] # idTree != modif

Data[idTree %in% idModif, idTree := duplicatedID] # we duplicate the idTree on the previous selected idTree  

Data[CensusYear == 2020 & idTree == duplicatedID] # to check
```


# Abnomal growth
# Abnomal recruit

# Save this test data in the package
```{r}
TestData <- Data
usethis::use_data(TestData)
```


