---
title: "RecruitmentCorrection"
author: "Vincyane Badouard"
date: "31/03/2022"
output: html_document
---

Inspiration: adaptation from Camille Piponiot's and Nino Page (ForestData package) original correction functions.

A utiliser après avoir corrigé le statut et la taille de l’arbre.

Return: Add rows for forgotten recruits with them estimated DBH in the 'DBHCor' column
+ create a 'CorrectedRecruit' col (logical)
+ and fill the 'Comment' column : "This DBH is/was the 1st recorded for this tree, according to its annual growth and the census done for this plot, it should have been recruited earlier according to your protocol (MinDBH)."

+ et col pour le type de cor ? si plsrs méthodes

Details :
If the size of the tree has never changed, or if there is only one value the same value is kept for the added forgotten recruits.

- If the 1st DBH is larger than it would have been if at the previous census it was at the minimum DBH :
DBH[1] > minDBH + (recrutement year - census before)* positivegrowththreshold  


DBH : diameter in cm

Tests :
- DataTree[CorrectedRecruit %in% TRUE, DBHCor] >= MinDBH
- if only 1 DBH value : DataTree[CorrectedRecruit %in% TRUE, DBHCor] == DataTree[CorrectedRecruit %in% FALSE, DBHCor]
- all(cresc < PositiveGrowthThreshold & cresc_abs > NegativeGrowthThreshold) # no aberrant growth

# Packages libraries
```{r, include = FALSE}
library(TreeData)
library(data.table)
library(ggplot2)
# importfrom stats::lm
```

# Import data & other inputs
```{r}
data("TestData") # import data
# INPUTS
## data.frame to data.table
setDT(TestData) # with "set" "<-" is not necessary

# Args
MinDBH = 10 # (numeric, 1 value) in cm : minimum diameter of trees inventoried according to your protocol
PositiveGrowthThreshold = 5 # (numeric, 1 value) in cm/year : a tree widening by more than x cm/year is considered abnormal
InvariantColumns = "Site" # Vector with the names of the columns that are supposed to have always the same value for each measurement of the same tree (character)
DetectOnly = FALSE # (logical)

```

# Create data with overgrown recruits
```{r}
DataTree <- data.table(IdTree = "a", # 4 ind
                       Year = seq(2000,2008, by = 2), # 9 DBH values each (2 years between 2 census)
                       DBHCor  = as.numeric(c(13:17)), # 1cm/census (0.5 cm/year), 1st DBH = 13 > MinDBH. In 1998 the tree was 12cm, 11cm in 1996 and 10cm in 1994
                       Site = "Imaginary forest"
)
DataTree
PlotCensuses = seq(1996,2016, by = 2)
```


# Ask to the user

# Check args
```{r}
# Data
if (!inherits(Data, c("data.table", "data.frame")))
  stop("Data must be a data.frame or data.table")

# MinDBH/PositiveGrowthThreshold (numeric, 1 value)
if(!all(unlist(lapply(list(MinDBH, PositiveGrowthThreshold), length)) %in% 1) |
   !all(unlist(lapply(list(MinDBH, PositiveGrowthThreshold),
                      inherits, c("numeric", "integer")))))
  stop("The 'MinDBH' and 'PositiveGrowthThreshold'' arguments
         of the 'RecruitmentCorrection' function must be 1 numeric value each")

# InvariantColumns
if (!inherits(InvariantColumns, "character"))
  stop("'InvariantColumns' argument must be of character class")

# DetectOnly (logical)
if(!all(unlist(lapply(list(DetectOnly),
                      inherits, "logical"))))
  stop("The 'DetectOnly' argument
         of the 'RecruitmentCorrection' function must be logicals")

# DBH column exists
if(!"DBH" %in% names(Data))
  stop("The 'DBH' (Diameter at Breast Height) column does't exist in the dataset")

# Check if the InvariantColumns name exists in DataTree
for(c in InvariantColumns){
  if(!c %in% names(DataTree)){
    stop(paste("InvariantColumns argument must contain one or several column names (see help)."
               ,c,"is apparently not a dataset's column"))
  }
}

```

# Correction
if the 'DBHCor' col doesnt exist create from the 'DBH' col ?
[Nino ne corrige que si l'arbre a déjà été vivant]

# Compute absolute and annual diameter incrementation
```{r}
DBHCor <- DataTree[,DBHCor]
Year <- DataTree[,Year]
# Initialisation
cresc <- cresc_abs <- rep(0, length(DBHCor) - 1) # (cresc[1] corresponds to the 2nd DBH)
# cresc_abs : absolute diameter increment (not divided by time between census)
# cresc : annual diameter increment

if (sum(!is.na(DBHCor)) > 1) { # if there is at least 1 measurement
  
  # Annual diameter increment (cresc)
  cresc[which(!is.na(DBHCor))[-1] - 1] <- # 8 cresc for 9 dbh values ([-1]), shift all indices by 1 to the left (-1)
    diff(DBHCor[!is.na(DBHCor)]) / diff(Year[!is.na(DBHCor)]) # DBH difference between pairwise censuses / time difference between pairwise censuses
  
  # Absolute diameter increment (not divided by time between census) (cresc_abs)
  cresc_abs[which(!is.na(DBHCor))[-1] - 1] <- diff(DBHCor[!is.na(DBHCor)]) 
  
}
cresc_abs
cresc

```

```{r}
# RecruitYear = 2000

PlotCensuses <- sort(PlotCensuses, decreasing = FALSE) # order years in the vector
#Make sure that it is ordered by increasing census time
DataTree <- DataTree[order(Year)] # order years in the data


if(any(!is.na(LifeStatusCor) & DataTree$LifeStatusCor == 1)){ # already been alive tree
  
  RecruitYear <- min(DataTree$Year[!is.na(DataTree$LifeStatusCor) & DataTree$LifeStatusCor == 1]) # recruitment year
  
  if(min(DataTree$Year) > min(PlotCensuses) & sum(!is.na(DataTree$DBHCor)) > 0){ # if the 1st census of the plot is smaller than the 1st census of the tree, and there are measured DBH
    
    PrevCens <- PlotCensuses[which(PlotCensuses == RecruitYear)-1] # 1 census before the recruit year among the plot censuses
    
    #### If the 1st DBH is larger than it would have been if at the previous census it was at the minimum DBH ####
    FirstDBH <- DataTree[!is.na(DBHCor), DBHCor][1] # 1st measured DBH
    
    # if(DataTree$DBHCor[1] > (MinDBH + (RecruitYear - PrevCens) * PositiveGrowthThreshold)){ # ah ben il detecte pas mes oublis pcq il considère l'erreur que si l'écart est superieur à la limite de croissance
    if(FirstDBH > (MinDBH + (RecruitYear - PrevCens) * cresc[1])){ # ma proposition
      
      MissingCens <- PlotCensuses[which(PlotCensuses < RecruitYear)] # the previous missing censuses 
      
      #### Create new rows for the forgotten recruits ####
      if(length(MissingCens) > 0){
        if("Plot" %in% names(DataTree)){ # if we have the plot info
          NewRow <- data.table(IdTree = unique(DataTree$IdTree), # the IdTree
                               Year = NA, # the censuses to add
                               Plot = unique(DataTree$Plot), # the unique plot of the tree
                               CorrectedRecruit = TRUE, # there are corrected recruits
                               stringsAsFactors = FALSE) # do not convert characters into factors
        }
        else{
          NewRow <- data.table(IdTree = unique(DataTree$IdTree), # the IdTree
                               Year = NA, # the censuses to add
                               CorrectedRecruit = TRUE, # there are corrected recruits
                               stringsAsFactors = FALSE) # do not convert characters into factors
        }
        
        if(length(InvariantColumns) > 0){ # if there are invariant columns
          
          NewRow[,(InvariantColumns) := NA] # empty the invariant columns for the added rows
          
          # Fill in the invariant columns in the added rows
          NewRow <- FillinInvariantColumns(NewRow = NewRow,
                                           InvariantColumns = InvariantColumns,
                                           DataTree = DataTree,
                                           IdTree = unique(DataTree$IdTree))
        }
        
        # Multiply this new row the number of times as well as the number of absents
        NewRows <- do.call("rbind", replicate(n = length(MissingCens), NewRow, simplify = FALSE))
        NewRows[, Year := MissingCens]
        
        # Add these rows in the dataset
        DataTree <- rbindlist(list(DataTree, NewRows), use.names=TRUE, fill=TRUE)
        DataTree[, CorrectedRecruit := ifelse(is.na(CorrectedRecruit), FALSE, CorrectedRecruit)] # FALSE for the other rows
        
        DataTree <- DataTree[order(Year)] # order by Year
        
        #### Linear regression (DBH ~ Year) ####
        coef <- stats::lm(
          DataTree[!is.na(DBHCor), DBHCor] ~ DataTree[!is.na(DBHCor), Year])$coefficients
        
        if(is.na(coef[2])) { # if no slope
          ### if only 1 DBH value: replace all non-recruited DBH by this value
          DataTree[Year < RecruitYear, DBHCor := coef[1]]
        }
        else {
          
          # Estimate the recruits DBHCor with linear extrapolation
          RecruitsDBH <- coef[1] + DataTree[Year < RecruitYear, Year]*coef[2] # y = b + ax. Min entre ces DBH inférés et le 1er DBH
          
          # If estimated DBHCors are higher than the first measured DBHCors, (comment c possible ?)
          # these are replaced by first measured DBHCors.
          for(y in 1: length(RecruitsDBH)){
            RecruitsDBH[y] <- min(RecruitsDBH[y], FirstDBH)
          }
          DataTree[Year < RecruitYear, ("DBHCor") := RecruitsDBH]
        }
        
        # UselessRows: added trees under the MinDBH
        UselessRows <- ((DataTree[, DBHCor] < MinDBH) & (DataTree[, Year] %in% MissingCens))
        
        if(any(UselessRows)){
          DataTree <- DataTree[-which(UselessRows)] # remove them
        }
        
        
      } # end: missing censuses to add
    } # end: overgrown recruit
  } # end: if the plot have previous censuses
} # end: ony for alive tree
```

