---
title: "StatusCorrection"
author: "Vincyane Badouard"
date: "16/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Inspiration: ForestData::correct_alive() and .correct_alive_tree()

Return: Create the 'LifeStatusCor' column

- if usesize : if DBH != NA -> Alive
- [*Dead* > Alive] -> NA
- add rows for the forgotten censuses between 2 'Alive'
- [Alive > *Dead*/*NA* > Alive] -> Alive
- [Alive > *NA* > Dead] -> NA
- [Alive > *Dead* > NA] -> Dead

- [Alive > *NA* > *NA*]:
if deathconfirmation > nrow(LifeStatus = "NA") -> NA
if deathconfirmation =< nrow(LifeStatus = "NA") -> Dead


# Packages libraries
```{r, include = FALSE}
library(TreeData)
library(data.table)
```


# Import data & other inputs
```{r}
data("TestData") # import data
# INPUTS
Data = TestData
## data.frame to data.table
setDT(Data) # with "set" "<-" is not necessary

# Args
Censuses = c(2011:2021) # censuses for the plot in which the tree is (numeric)
deathconfirmation = 2 # (numeric)
usesize = FALSE # (logical)
DetectOnly = FALSE # (logical)
InvariantColumns = c("Site",
                     "Genus",
                     "Species",
                     "Family",
                     "ScientificName") # user choice

# Supply names of columns that have 0 variance (= invariant columns)
# vari <- lapply(DataTree, var)
# df <- data.frame(InvariantC = c(vari == 0))
# InvariantColumns <- setDT(df, keep.rownames = TRUE)[InvariantC == TRUE| is.na(InvariantC), rn] # detection
# 
# InvariantColumns <- InvariantColumns[!InvariantColumns %in% c("CensusDate", "CensusDateCertainty", "LifeStatus", "MeasCode", "CorrCode")]  
```

# Ask to the user
```{r}
# The LifeStatus coding
# here
# 1 = alive (TRUE)
# 0 = dead (FALSE)
# NA = unseen

RemoveRAfterDeath = FALSE # "After correction do you want to delete the rows concerning the tree after its death?" (logical)
```

# Function per tree (IdTree)
## Data creation
```{r}
# Wanted seq: Dead > Alive > NA > Dead > Alive > NA > NA > Dead > NA

DataTree <- Data[IdTree == "100658"]

AddR <- DataTree[2:5] # the rows to copy
AddR[, CensusYear := c(2012:2015)] # the rows to add

DataTree <- rbindlist(list(DataTree, AddR)) # add rows

DataTree <- DataTree[order(CensusYear)]

DataTree[, LifeStatus := c(FALSE, TRUE, NA, FALSE, TRUE, NA, NA, FALSE, NA)] # write the sequence

```

## Code
```{r}
# Arrange year in ascending order
DataTree <- DataTree[order(CensusYear)] # order de dt


if(!length(unique(DataTree$Plot)) == 1){ # if there are several plots for the same idtree
  stop(paste0("tree ",unique(DataTree$IdTree)," has multiple plots: " ,paste0(unique(DataTree$Plot), collapse = "/")))
}
```

### Use the size presence as a witness of the living status of the tree
```{r}
if(usesize){
  
  DataTree <- GenerateComment(DataTree,
                              condition = !is.na(DataTree[, DBH]) &
                                !DataTree[,LifeStatus] %in% TRUE,
                              comment = "A measured tree is a living tree")
  
  
  if(DetectOnly %in% FALSE){
    DataTree[!is.na(DBH), LifeStatusCor := TRUE] # c'est tout ? une taille = vivant ? 
  }
  
}else{
  
  if(DetectOnly %in% FALSE){
    DataTree[, LifeStatusCor := LifeStatus] # we will work on a new col and keep the original col intact
  }
}
```

### Sequence analyse
```{r}
# The first tree life status in the dataset
FirstStatus <- DataTree[CensusYear == min(CensusYear), LifeStatus]

# The first and last Alive record year
FirstAliveYear <- min(DataTree[LifeStatus %in% TRUE, CensusYear])
LastAliveYear <- max(DataTree[LifeStatus %in% TRUE, CensusYear])

# Before/between positions (rows id)
FirstAlive <- which(DataTree$LifeStatus %in% TRUE)[1] # the 1st seen alive
LastAlive <-  max(which(DataTree$LifeStatus == 1)) # the last seen alive

Before <- which(DataTree$CensusYear < FirstAliveYear) # Before the first alive
Between <- which(DataTree$CensusYear > FirstAliveYear & DataTree$CensusYear < LastAliveYear) # Between the fist and last alive

```

### Absents (logical vector of the Censuses length)
```{r}
# if tree has ever been recorded dead
if(any(DataTree$LifeStatus %in% FALSE)){
  # The last time where the tree has been recorded dead (in case there are several)
  LastDeathRecord <- max(DataTree[LifeStatus %in% FALSE, CensusYear])
  
  After <- which(DataTree$CensusYear > LastDeathRecord) # After the last death record
  
  # If there is any "Alive" report after last reported death
  if(any(DataTree$LifeStatus[After] %in% TRUE)) {
    # Absents are the absent record years among the plot censuses from the 1st alive record
    Absents <- (Censuses > FirstAliveYear & !Censuses %in% DataTree$CensusYear)
    
  }else{
    # Absents are the absent record years between first alive record and the last death record
    Absents <- (Censuses > FirstAliveYear &
                  Censuses < LastDeathRecord & # death = the end 
                  !Censuses %in% DataTree$CensusYear)
  }
  
}else{ # if tree has not been reported dead yet
  
  # Absents are the absent record years among the plot censuses from the 1st alive record
  Absents <- (Censuses > FirstAliveYear & !Censuses %in% DataTree$CensusYear)
  
}

# En DetectOnly, je renvoie qqchose ? quoi ? ########

```

### Function to fill in the invariant columns in the added rows
```{r}
FillinInvariantColumns <- function(NewRow, InvariantColumns, DataTree, IdTree){
  
  for(j in InvariantColumns){
    
    if(any(is.na(NewRow[,get(j)]))){ # if the column is empty in the new rows (the "absent" trees)
      
      uni <- unique(DataTree[, get(j)])
      
      if(length(uni) > 1){ # if the "invariant column is not invariant
        stop("The variable ",
             j,
             " that you defined as a non-varying column -i.e. supposed to have always the same value for each measurement of the same tree- has multiple values for tree ",
             IdTree,
             " and takes the values ",
             uni)
      }
      else if(length(uni) == 0){ # no value in the invariant column
        stop("The variable ",j," has no value for individual ",IdTree)
      }
      else{
        NewRow[is.na(get(j)), (j) := uni] # fill the invariant column in NewRow with their (unique) value
      }
    }
  }
  
  return(NewRow)
}

```

### Creating rows for absents
```{r}
if(DetectOnly %in% FALSE){
  
  Nabs <- sum(Absents) # absent is a logical vector giving the census times for which trees were not seen.
  
  if(Nabs > 0){ # if there are absents
    # if(DataTree$Plot[1] == 1) print(DataTree$Plot[1])
    NewRow <- data.table(IdTree = unique(DataTree$IdTree),     # the idtree
                         CensusYear = Censuses[Absents],       # absent censuses list
                         LifeStatus = NA,                    # not seen
                         LifeStatusCor = NA,               # no corrected status for now
                         Plot = unique(DataTree$Plot),  # the unique plot in DataTree
                         SubPlot = unique(DataTree$SubPlot),  # the unique subplot in DataTree
                         stringsAsFactors =  FALSE)      # do not convert characters into factors
    NewNames <- names(NewRow)
    
    if(length(InvariantColumns) > 0){ # if there are invariant columns
      
      NewRow[,(InvariantColumns) := NA] # empty the invariant columns for the added rows
      
      # Fill in the invariant columns in the added rows
      NewRow <- FillinInvariantColumns(NewRow = NewRow,
                                       InvariantColumns = InvariantColumns,
                                       DataTree = DataTree,
                                       IdTree = unique(DataTree$IdTree))
    }
    
    # Multiply this new row the number of times as well as the number of absents
    NewRows <- do.call("rbind", replicate(n = Nabs, NewRow, simplify = FALSE))
    
    # Add these rows in the dataset
    DataTree <- rbindlist(list(DataTree, NewRows), use.names=TRUE, fill=TRUE)
    
    DataTree <- DataTree[order(CensusYear)] # order by time
    
  } # end: Nabsents > 0
}

```


### Alive > *Alive* > Alive
```{r}
DataTree <- GenerateComment(DataTree,
                            condition = seq.int(nrow(DataTree)) %in% (FirstAlive:LastAlive) &
                              !DataTree[, LifeStatus] %in% TRUE,
                            comment = "Between 2 alive occurrences of the tree, it cannot be dead")

if(DetectOnly %in% FALSE){
  DataTree[FirstAlive:LastAlive, LifeStatusCor := TRUE] # so all between is alive
}

```

### Enough/not enough occurrences of death to validate it (arrétée ici)
```{r}
if(LastAlive != nrow(DataTree)){ # if the last seen alive is not the last row of the database
  
  # if the one after the last one seen alive is dead:
  if(DataTree[LastAlive +1, LifeStatus] %in% FALSE){
    
    # Remove rows after the death (after correction) (User choice)
    if(RemoveRAfterDeath %in% TRUE)
      
      DataTree <- DataTree[-(LastAlive +2:nrow(DataTree)),]
  }
  else{
    # enough occurrences of death to validate it
    if(LastAlive < nrow(DataTree)-(dead_confirmation_censuses -1)){ # -1 to count the last row and check if the data finish with at least 2 deaths
      DataTree$LifeStatusCor[LastAlive +1] <- 0 # validated: the one after the last one seen alive is dead
      DataTree <- DataTree[-(LastAlive +2:nrow(DataTree)),] # not understand
    }
    else{ # there are not enough occurrences of death to validate it
      DataTree[(LastAlive +1):nrow(DataTree),"LifeStatusCor"] <- NA # the death is not sure
    }
  }
}
```

### Before the first alive 
```{r}
if(first_seen_alive > 1){ # if the first seen alive is not the first row
  DataTree <- DataTree[-(1:(first_seen_alive-1)),] # remove from the first row to to the last before the first seen alive
}
```

