---
title: "StatusCorrection"
author: "Vincyane Badouard"
date: "16/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Inspiration: ForestData::correct_alive() and .correct_alive_tree()

Return: Create the 'LifeStatusCor' column and fill the 'Comment' column

- if UseSize : if DBH != NA -> Alive
- [*Dead* > Alive] -> NA
- add rows for the forgotten censuses between 2 'Alive'
- [Alive > *Dead*/*NA* > Alive] -> Alive
- [Alive > *NA* > Dead] -> NA
- [Alive > *Dead* > NA] -> Dead

- [Alive > *NA* > *NA*]:
if DeathConfirmation > unseens -> NA
if DeathConfirmation =< unseens -> Dead


# Packages libraries
```{r, include = FALSE}
library(TreeData)
library(data.table)
```

# Import data & other inputs
```{r}
data("TestData") # import data
# INPUTS
Data = TestData
## data.frame to data.table
setDT(Data) # with "set" "<-" is not necessary

# Args
Censuses = c(2011:2021) # censuses for the plot in which the tree is (numeric)
DeathConfirmation = 2 # (numeric)
UseSize = FALSE # (logical)
DetectOnly = FALSE # (logical)

RemoveRAfterDeath = FALSE # "After correction do you want to delete the rows concerning the tree after its death?" (logical)
RemoveRBeforeAlive = FALSE # "Do you want to delete the rows about the tree before it was seen alive for the 1st time?" (logical)

InvariantColumns = c("Site",
                     "Genus",
                     "Species",
                     "Family",
                     "ScientificName") # user choice

# Supply names of columns that have 0 variance (= invariant columns)
# vari <- lapply(DataTree, var)
# df <- data.frame(InvariantC = c(vari == 0))
# InvariantColumns <- setDT(df, keep.rownames = TRUE)[InvariantC == TRUE| is.na(InvariantC), rn] # detection
# 
# InvariantColumns <- InvariantColumns[!InvariantColumns %in% c("CensusDate", "CensusDateCertainty", "LifeStatus", "MeasCode", "CorrCode")]  
```

# Ask to the user
```{r}
# The LifeStatus coding
# here
# 1 = alive (TRUE)
# 0 = dead (FALSE)
# NA = unseen
```

# Function per tree (IdTree)
## Data creation
```{r}
# Wanted seq: Dead > Alive > NA > Dead > Alive > NA > NA > Dead > NA

DataTree <- Data[IdTree == "101623"] # For all deads : 100635

AddR <- DataTree[2:5] # the rows to copy
AddR[, CensusYear := c(2012:2015)] # the rows to add

DataTree <- rbindlist(list(DataTree, AddR)) # add rows

DataTree <- DataTree[order(CensusYear)]

DataTree[, LifeStatus := c(FALSE, TRUE, NA, FALSE, TRUE, NA, NA, FALSE, NA)] # write the sequence

```

## Code
```{r}
# Arrange years in ascending order
DataTree <- DataTree[order(CensusYear)] # order de dt


if(!length(unique(DataTree$Plot)) == 1){ # if there are several plots for the same idtree
  stop(paste0("tree ",unique(DataTree$IdTree)," has multiple plots: " ,paste0(unique(DataTree$Plot), collapse = "/")))
}
```

### Use the size presence as a witness of the living status of the tree
```{r}
if(UseSize){
  
  DataTree <- GenerateComment(DataTree,
                              condition = !is.na(DataTree[, DBH]) &
                                !DataTree[,LifeStatus] %in% TRUE,
                              comment = "A measured tree is a living tree")
  
  
  if(DetectOnly %in% FALSE){
    DataTree[!is.na(DBH), LifeStatusCor := TRUE] # c'est tout ? une taille = vivant ? non ya encore des corrections possibles !!
  }
  
}else{
  
  if(DetectOnly %in% FALSE){
    DataTree[, LifeStatusCor := LifeStatus] # we will work on a new col and keep the original col intact
  }
}
```

### Sequence analyse
```{r}
# if tree has ever been recorded alive
if(any(DataTree$LifeStatusCor %in% TRUE)){
  # The first tree life status in the dataset
  # FirstStatus <- DataTree[CensusYear == min(CensusYear), LifeStatusCor]
  
  # The first and last Alive record year
  FirstAliveYear <- min(DataTree[LifeStatusCor %in% TRUE, CensusYear], na.rm = TRUE)
  # LastAliveYear <- max(DataTree[LifeStatusCor %in% TRUE, CensusYear], na.rm = TRUE)
  
  # Before/between positions (rows id)
  FirstAlive <- which(DataTree$LifeStatusCor %in% TRUE)[1] # the 1st seen alive
  LastAlive <-  max(which(DataTree$LifeStatusCor %in% TRUE)) # the last seen alive

  # Before <- which(DataTree$CensusYear < FirstAliveYear) # Before the first alive
  # Between <- which(DataTree$CensusYear > FirstAliveYear & DataTree$CensusYear < LastAliveYear) # Between the fist and last alive
}
```

### Absents (logical vector of the Censuses length)
Lignes manquantes :
- pendant la vie
- entre la vie et la mort
- entre le 1er et le dernier census (quand pas de vivant)
```{r}
if(any(DataTree$LifeStatusCor %in% TRUE)){
  
# if tree has ever been recorded dead
if(any(DataTree$LifeStatusCor %in% FALSE)){
  # The last time where the tree has been recorded dead (in case there are several)
  LastDeathRecord <- max(DataTree[LifeStatusCor %in% FALSE, CensusYear])
  
  After <- which(DataTree$CensusYear > LastDeathRecord) # After the last death record
  
  # If there is any "Alive" report after last reported death
  if(any(DataTree$LifeStatusCor[After] %in% TRUE)) {
    # Absents are the absent record years among the plot censuses from the 1st alive record
    Absents <- (Censuses > FirstAliveYear & !Censuses %in% DataTree$CensusYear)
    
  }else{
    # Absents are the absent record years between first alive record and the last death record
    Absents <- (Censuses > FirstAliveYear &
                  Censuses < LastDeathRecord & # death = the end 
                  !Censuses %in% DataTree$CensusYear)
  }
  
}else{ # if tree has not been reported dead yet
  
  # Absents are the absent record years among the plot censuses from the 1st alive record
  Absents <- (Censuses > FirstAliveYear & !Censuses %in% DataTree$CensusYear)
  
}

# if no one alive
}else{

  # La j'ai choisi de ne rajouter les lignes absentes qu'entre le census min et max de l'arbre
  # Si tout est FALSE effectivement ça ne sert à rien de rajouter des lignes apres, mais des lignes avant ? il risque d'en avoir beaucoup, et on ne pourra mettre qu'NA
  # Pour tout est NA, ça aurait un intéret de rajouter des lignes avant-après ?
Absents <- (Censuses > min(DataTree$CensusYear, na.rm = TRUE) & # entre les bornes, pas avnt pas après
              Censuses < max(DataTree$CensusYear, na.rm = TRUE) &
              !Censuses %in% DataTree$CensusYear)

}

# En DetectOnly, je renvoie qqchose ? quoi ? ########

```

### Function to fill in the invariant columns in the added rows
```{r}
FillinInvariantColumns <- function(NewRow, InvariantColumns, DataTree, IdTree){
  
  for(j in InvariantColumns){
    
    if(any(is.na(NewRow[,get(j)]))){ # if the column is empty in the new rows (the "absent" trees)
      
      uni <- unique(DataTree[, get(j)])
      
      if(length(uni) > 1){ # if the "invariant column is not invariant
        stop("The variable ",
             j,
             " that you defined as a non-varying column -i.e. supposed to have always the same value for each measurement of the same tree- has multiple values for tree ",
             IdTree,
             " and takes the values ",
             uni)
      }
      else if(length(uni) == 0){ # no value in the invariant column
        stop("The variable ",j," has no value for individual ",IdTree)
      }
      else{
        NewRow[is.na(get(j)), (j) := uni] # fill the invariant columns in NewRow with their (unique) value
      }
    }
  }
  
  return(NewRow)
}

```

### Creating rows for absents
```{r}
if(DetectOnly %in% FALSE){
  
  Nabs <- sum(Absents) # absent is a logical vector giving the census times for which trees were not seen.
  
  if(Nabs > 0){ # if there are absents
    # if(DataTree$Plot[1] == 1) print(DataTree$Plot[1])
    NewRow <- data.table(IdTree = unique(DataTree$IdTree),     # the idtree
                         CensusYear = Censuses[Absents],       # absent censuses list
                         LifeStatus = NA,                    # not seen
                         LifeStatusCor = NA,               # no corrected status for now
                         Plot = unique(DataTree$Plot),  # the unique plot in DataTree
                         SubPlot = unique(DataTree$SubPlot),  # the unique subplot in DataTree
                         stringsAsFactors =  FALSE)      # do not convert characters into factors
    NewNames <- names(NewRow)
    
    if(length(InvariantColumns) > 0){ # if there are invariant columns
      
      NewRow[,(InvariantColumns) := NA] # empty the invariant columns for the added rows
      
      # Fill in the invariant columns in the added rows
      NewRow <- FillinInvariantColumns(NewRow = NewRow,
                                       InvariantColumns = InvariantColumns,
                                       DataTree = DataTree,
                                       IdTree = unique(DataTree$IdTree))
    }
    
    # Multiply this new row the number of times as well as the number of absents
    NewRows <- do.call("rbind", replicate(n = Nabs, NewRow, simplify = FALSE))
    
    # Add these rows in the dataset
    DataTree <- rbindlist(list(DataTree, NewRows), use.names=TRUE, fill=TRUE)
    
    DataTree <- DataTree[order(CensusYear)] # order by time
    
  } # end: Nabsents > 0
}

```


### Alive > *Alive* > Alive
```{r}
if(any(DataTree$LifeStatusCor %in% TRUE)){
  
  DataTree <- GenerateComment(DataTree,
                              condition = seq.int(nrow(DataTree)) %in% (FirstAlive:LastAlive) &
                                !DataTree[, LifeStatusCor] %in% TRUE,
                              comment = "Between 2 alive occurrences of the tree, it cannot be dead")
  
  if(DetectOnly %in% FALSE){
    DataTree[FirstAlive:LastAlive, LifeStatusCor := TRUE] # so all between is alive
  }
}
```

### Enough/not enough occurrences of death to validate it
S'il ya des NA dans la sequence après le dernier Alive, s'ils sont en nbr >/= DeathConfirmation -> Dead
On ne peut être mort qu'après la vie.
```{r}
if(any(DataTree$LifeStatusCor %in% NA)){
  if(any(DataTree$LifeStatusCor %in% TRUE)){
    
    
    # If there are things after the last occurrence of life
    if(LastAlive != nrow(DataTree)){ # if the last seen alive is not the last row of the database
      
      #### if the one after the last one seen alive is Dead ####
      if(DataTree[LastAlive +1, LifeStatusCor] %in% FALSE){
        
        # Remove rows after the death (after correction) (User choice)
        if(RemoveRAfterDeath %in% TRUE)
          DataTree <- DataTree[-(LastAlive +2:nrow(DataTree)),]
        
      }
      #### if the one after the last one seen alive is Unseen ####
      else if(DataTree[LastAlive +1, LifeStatusCor] %in% NA){
        
        ##### If there is still a "death" occurrence #####
        if(any(DataTree$LifeStatusCor %in% FALSE)){
          
          LastDeath <- max(which(DataTree$LifeStatusCor %in% FALSE))
          
          ###### If the death is not the last record ######
          if(LastDeath < nrow(DataTree)){
            unseen <- sum(DataTree[(LastAlive +1):(LastDeath-1), LifeStatusCor] %in% NA) # NA until the death (logicals vector) 
            
          }else{
            unseen <- sum(DataTree[(LastAlive +1):nrow(DataTree), LifeStatusCor] %in% NA) # NA until the dataset end (logicals vector) 
          }
          ##### No death record #####
        }else{
          unseen <- sum(DataTree[(LastAlive +1):nrow(DataTree), LifeStatusCor] %in% NA) # NA until the dataset end (logicals vector) 
        }
        
        if(DeathConfirmation <= unseen){
          
          # The comment
          DataTree <- GenerateComment(DataTree,
                                      condition = seq.int(nrow(DataTree)) %in% ((LastAlive +1):(LastAlive +unseen)) &
                                        DataTree[, LifeStatusCor] %in% NA,
                                      comment = "When the tree is unseen a number of times >= DeathConfirmation, it is considered dead")
          
          if(DetectOnly %in% FALSE){
            
            # The correction
            DataTree[(LastAlive +1):(LastAlive +unseen), LifeStatusCor := FALSE] # Death validated
            
            # Remove rows after the death (after correction) (User choice)
            if(RemoveRAfterDeath %in% TRUE)
              DataTree <- DataTree[-(LastAlive +2:nrow(DataTree)),]
            
          } # correction end
          
        } # else if(DeathConfirmation > unseen) NAs remain NAs
        
      }
    } # If there nothing after the last occurrence of life
  } # if there is any alive
} # any NA ?

```

### Before the first alive 
Qu'y a t'il avant le 1er vivant ? 
Mort ? impossibble -> NA
NA ? ben ça reste NA
```{r}
if(any(DataTree$LifeStatusCor %in% TRUE)){
  
  if(FirstAlive > 1){ # if the first seen alive is not the first row -> NA/dead
    
    # The comment
    DataTree <- GenerateComment(DataTree,
                                condition = seq.int(nrow(DataTree)) %in% (1:(FirstAlive -1)) &
                                  DataTree[, LifeStatusCor] %in% NA, # -> unseen
                                comment = "Isn't the tree alive?")
    
    DataTree <- GenerateComment(DataTree,
                                condition = seq.int(nrow(DataTree)) %in% (1:(FirstAlive -1)) &
                                  DataTree[, LifeStatusCor] %in% FALSE, # -> dead
                                comment = "Tree cannot be dead before being alive")
    
    
    if(DetectOnly %in% FALSE){
      
      # The correction (mettre un ifelse)
      DataTree[1:(FirstAlive -1), LifeStatusCor := ifelse(LifeStatusCor %in% FALSE, NA, LifeStatusCor)] # Tree cannot be dead before being alive -> NA
      
      # Remove rows after the death (after correction) (User choice)
      if(RemoveRBeforeAlive %in% TRUE)
        DataTree <- DataTree[-(1:(FirstAlive -1)),] # remove from the first to the last row before the first seen alive
      
    } # correction end
  }
}
```

### After the death
Quest-ce qu'il ya après la dernière mort ?
Alive impossible
Dead ben non c'est la dernière mort alors
NA ? apres la mort il n'ya que la mort.
```{r}
# If all rows count
if(RemoveRAfterDeath %in% FALSE){
  
  # If there is still a "death" occurrence
  if(any(DataTree$LifeStatusCor %in% FALSE)){
    
    LastDeath <- max(which(DataTree$LifeStatusCor %in% FALSE))
    
    # If the death is not the last record
    if(LastDeath < nrow(DataTree)){
      
      # The comment
      DataTree <- GenerateComment(DataTree,
                                  condition = seq.int(nrow(DataTree)) %in% ((LastDeath +1):nrow(DataTree)) &
                                    DataTree[, LifeStatusCor] %in% NA,
                                  comment = "After its death the tree is still dead")
      
      if(DetectOnly %in% FALSE){
        
        # The correction
        DataTree[(LastDeath +1):nrow(DataTree), LifeStatusCor := FALSE] # After death there is only death
        
      } # correction end
    } # the death isn't the last record
  } # there is still a death
} # we want all the deaths!
```

### Dead > *Dead* > Dead
S'il reste des NA entre 2 deads (après les vivants)
```{r}
if(any(DataTree$LifeStatusCor %in% FALSE)){


FirstDead <- which(DataTree$LifeStatusCor %in% FALSE)[1] # the 1st seen dead
LastDead <-  max(which(DataTree$LifeStatusCor %in% FALSE)) # the last seen dead

DataTree <- GenerateComment(DataTree,
                            condition = seq.int(nrow(DataTree)) %in% (FirstDead:LastDead) &
                              !DataTree[, LifeStatusCor] %in% FALSE,
                            comment = "Between 2 dead occurrences of the tree, it is dead")

if(DetectOnly %in% FALSE){
  DataTree[FirstDead:LastDead, LifeStatusCor := FALSE] # so all between is dead
}
}

```

### After the death 
```{r}
# If all rows count
if(RemoveRAfterDeath %in% FALSE){
  
  # If there is still a "death" occurrence
  if(any(DataTree$LifeStatus %in% FALSE)){
    
    LastDeath <- max(which(DataTree$LifeStatus %in% FALSE))
    
    # If the death is not the last record
    if(LastDeath < nrow(DataTree)){
      
      # The comment
      DataTree <- GenerateComment(DataTree,
                                  condition = seq.int(nrow(DataTree)) %in% ((LastDeath +1):nrow(DataTree)) &
                                    DataTree[, LifeStatus] %in% NA,
                                  comment = "After its death the tree is still dead")
      
      if(DetectOnly %in% FALSE){
        
        # The correction
        DataTree[(LastDeath +1):nrow(DataTree), LifeStatusCor := FALSE] # After death there is only death
        
      } # correction end
    } # the death isn't the last record
  } # there is still a death
} # we want all the deaths!
```

# For all the trees 
```{r}
c("101433","101435","101436","101437","101438","101439","101440",
  "101441","101442","101443","101444","101446","101447","101448")
```

```{r}
# order IdTrees and times in ascending order
Data <- Data[order(IdTree, CensusYear)]

# IdTre
Ids <- as.vector(na.omit(unique(Data$IdTree))) # Tree Ids

# i = "100635"
a <- do.call(rbind, lapply(Ids, function(i) StatusCorrectionByTree(
  Data[IdTree %in% i], # per IdTree, all censuses
  Censuses = as.vector(na.omit( # rm NA
    unique(Data[Plot %in% unique(Data[IdTree %in% i, Plot]),  CensusYear]) # the censuses for the plot in which the tree is
  )),
  UseSize = UseSize,
  DetectOnly = DetectOnly,
  
  RemoveRBeforeAlive = RemoveRBeforeAlive,
  RemoveRAfterDeath = RemoveRAfterDeath
)
)) # do.call apply the 'rbind' to the lapply result

```



